### 2017-10-16 paper.tex
wqCommiting first draft by @rcurtin

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

% Could this be a better title?
\title{A generic and fast C++ optimization framework}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Shikhar Bhardwaj \\
  \AND
  Yannis Mentekidis \\
  \AND
  Marcus Edel \\
  \AND
  Ryan R. Curtin \\
  Symantec Corporation \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
The development of the {\bf mlpack} C++ machine learning library
(\url{http://www.mlpack.org/}) has required the design and implementation of a
flexible, robust optimization system that is able to solve the types of
arbitrary optimization problems that may arise all throughout machine learning
problems.  You care because:

\begin{itemize}
  \item It's easy to implement new optimizers.
  \item It's easy to implement new functions to be optimized.
  \item This system is more comprehensive than other techniques that may focus
only on neural networks or single problems.
  \item This system is also more specific and offers more room for optimization
than super general systems like scipy.optimize().
\end{itemize}

This abstract will need to be rewritten.
\end{abstract}

\section{Introduction}

In recent times, machine learning conferences have focused more and more on
optimization techniques.  For instance, this year at ICML (ICML 2017), every
single session had at least one track devoted solely to optimization
techniques.
<TODO: add some more filler here.  Show that optimization is an important part
of machine learning, in whatever ways.>

\begin{itemize}
  \item This sequence of events (or situation) motivates the development of a
generic, flexible, and fast optimization framework.
  \item We have implemented exactly this in {\bf mlpack}, and it's good.
\end{itemize}

{\bf <TODO: elaborate on above, add a few more sentences.>}

{\bf <TODO: mention related work, or really that we haven't found other
libraries that have a comprehensive infrastructure like the one we have built,
and that what we have built is more flexible and generic}

\section{mlpack: a scalable C++ machine learning library}

\begin{itemize}
  \item Some quick background on what mlpack is and what its goals are.
Probably the list of four goals is good here (though maybe not in itemized
format).
  \item Discuss why we've chosen C++, and how policy-based design gives us some
really nice flexibility.  (Maybe there is not fully space for that.)
\end{itemize}

\section{Requirements for optimizers and functions}

In general, we want to be able to consider any solver of the problem

\begin{equation}
\operatorname{argmin}_{x} f(x)
\end{equation}

\noindent for any function $f(x)$ that takes some vector input $x$.  However, it
is impossible to implement something both so generic and fast---for instance,
gradient-based approaches converge far more quickly than gradient-free
approaches (in general), so we must design an abstraction that is able to
simultaneously generalize to many problem types, as well as able to take
advantage of accelerations and optimizations.

Let us split the class of functions to be optimized into some subclasses:

\begin{itemize}
  \item {\bf arbitrary}: no assumptions can be made on $f(x)$
  \item {\bf differentiable}: $f(x)$ has a computable gradient $f'(x)$
  \item {\bf separable}: $f(x)$ is a sum of individual components: $f(x) =
\sum_{i} f_i(x)$
  \item {\bf categorical}: $x$ contains elements that can only take discrete
values
  \item {\bf numeric}: all elements of $x$ take values in $\mathcal{R}$
  \item {\bf sparse}: the gradient $f'(x)$ or $f_i(x)$ (for a separable
function) is sparse
  \item {\bf partially differentiable}: the gradient $f'_j(x)$ is computable for
individual elements $x_j$ of $x$
  \item {\bf bounded}: $x$ is limited in the values that it can take
\end{itemize}

Needless to say, it is impossible to create a system where every optimizer can
work with every possible type of function: a gradient-based optimizer cannot
reasonably be expected to work with an arbitrary function $f(x)$ where the
gradient is not computable or available.

Instead, the best we can hope to achieve is to maximize the flexibility
available, so that a user can easily implement a function $f(x)$ and have it
work with as many optimizers as possible.  For this, C++ policy-based design
aids us greatly: when implementing a function to be optimized, a user can
implement only a few methods and we can use C++ template metaprogramming to
check that the given functions match the requirements of the optimizer that is
being used.  When implementing an optimizer, a user can assume that the given
function to be optimized meets the required assumptions of the optimizers, and
encode those requirements.  Since we are using templates and C++, the resulting
code generated by the compiler can be identical to what a developer would write
if they were writing an optimizer specifically for the function $f(x)$ at
hand---this can provide significant efficiency gains.

{\bf <TODO: maybe we should add a table with optimizer types that work with
certain assumptions.  Maybe this can be a table, maybe a diagram, but some kind
of visual aid to split the space of objective function types would be nice.>}

\section{{\tt FunctionType} API}

In order to facilitate consistent implementations, we have defined a {\tt
FunctionType} API that describes all the functions that an objective function
may (or must) implement, depending on the subclass of objective function.

{\bf <TODO: we can add a list of each function and what it does.  A basic list
is below.  We can elaborate on this significantly.>}

\begin{itemize}
  \item {\tt Evaluate()}
  \item {\tt Gradient()}
  \item {\tt PartialGradient()} 
  \item ...
  \item constraints?
  \item discrete variable support?
  \item batch support
\end{itemize}

A user may implement as many (or as few) of these functions as they would like,
and this will allow them to use a larger set of optimizers.

{\bf <TODO: add very simple example function>}

\section{Static checking of functions}

{\bf <TODO: probably a better section title>}

Unfortunately, template metaprogramming can result in some very lengthy error
messages.  Therefore, we must be careful to ensure that a user is able to easily
debug a problem when they implement an objective function or gradient.  To
provide good error message output, we can use C++'s template metaprogramming
support to determine what methods a type has available.  Similarly, we can also
use static compile-time constants to denote the methods that are require by a
specific optimizer.  This is implemented via SFINAE \cite{sfinae}.  A {\tt
static\_assert()} is raised when a given objective function does not have the
required methods for a given optimizer.  {\bf <TODO: implement this, so we are
not lying.  But it should be easy enough, maybe an hour of work max.>}

For instance, when attempting to use the L-BFGS optimizer without having a
{\tt Gradient()} function implemented, the user will receive the following error
message:

{\bf <TODO: put error message here.  If we are short on space, let's cut this
example and make this section shorter.>}

\section{Supported optimizers and functions in mlpack}

Thanks to the easy abstraction, we have been able to provide support for a large
set of diverse optimizers and objective functions.  Below is a list of what is
currently available.

\begin{itemize}
  \item {\bf <TODO: add all optimizers>}
  \item {\bf <TODO: add all things that we can optimize>}
\end{itemize}

\section{Conclusion}

We have identified that the support for generic and robust optimization is not
currently available in most machine learning toolkits, and acted upon this
observation to provide an easy framework for both implementing new optimizers
and new objective functions to be optimized inside of the {\bf mlpack} machine
learning library.  This supports a

\end{document}

### 2017-10-17 paper.tex
Add example comparison table.

@@
 \usepackage{amsthm}
+\usepackage{array}
+\usepackage{booktabs}
+\usepackage{graphicx}
+\usepackage{wasysym}
+\usepackage{tabularx}
+\usepackage{threeparttable}
 
@@
 
-\usepackage{nips_2017}
+\usepackage[]{nips_2017}
 
@@
 \usepackage{microtype}      % microtypography
+
+\newcommand*\rot{\rotatebox{90}}
+\newcommand*\OK{\ding{51}}
 
@@
   Marcus Edel \\
+  Free University of Berlin \\
+  Arnimallee 7, 14195 Berlin \\
+  \texttt{marcus.edel@fu-berlin.de} \\
   \AND
@@
 and that what we have built is more flexible and generic}
+
+% Just to get a feeling of the look; Also, we could also just use \OK instead of \CURCLE or \LEFTcircle
+\begin{table}
+\centering
+    \begin{tabular}{@{} cl*{4}c @{}}
+        & & \multicolumn{4}{c}{} \\[2ex]
+        & & \rot{has framework}
+          & \rot{\shortstack[l]{problem-independent\\constraints}}
+          & \rot{supports x}
+          & \rot{supports x} \\
+        \cmidrule{2-6}
+        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE  \\
+        & vowpal wabbit      &  - & - & -  & -  \\
+        & TensorFlow         &  - & -  & -  & -  \\
+        & scikit-learn       &  - & -  & -  & - \\
+        & scipy              &  - & \LEFTcircle  & \LEFTcircle  & \LEFTcircle \\
+        \cmidrule[1pt]{2-6}
+    \end{tabular}
+    \begin{tablenotes}\footnotesize
+\item[*] Library feature comparison: \CIRCLE=\text{provides feature}, \LEFTcircle=\text{partially provides feature}, -=\text{does not provide feature}
+\end{tablenotes}
+\end{table}
 

### 2017-10-17 paper.tex
Add first optimizer.
Refactor introduction, fill in blank pieces.
I'm ok with this; maybe it can be improved but this is at least an ok start.

@@
 
-In recent times, machine learning conferences have focused more and more on
-optimization techniques.  For instance, this year at ICML (ICML 2017), every
-single session had at least one track devoted solely to optimization
-techniques.
-<TODO: add some more filler here.  Show that optimization is an important part
-of machine learning, in whatever ways.>
+Machine learning is a field that is inextricably intertwined with the field of
+optimization.  Countless machine learning techniques depend on the optimization
+of a given objective function; for instance, classifiers such as logistic
+regression \cite{cox1958regression}, metric learning methods like NCA
+\cite{goldberger2005neighbourhood}, manifold learning algorithms like MVU
+\cite{weinberger2006introduction}, and the extremely popular field of deep
+learning \cite{schmidhuber2015deep}.  Thanks to the attention focused on these
+problems, it is increasingly important in the field to have fast, practical
+optimizers.  This can explain the current focus on optimization: this year at
+ICML (2017), every single session had at least one track devoted solely to
+optimization techniques.
 
-\begin{itemize}
-  \item This sequence of events (or situation) motivates the development of a
-generic, flexible, and fast optimization framework.
-  \item We have implemented exactly this in {\bf mlpack}, and it's good.
-\end{itemize}
+Therefore, the need is real to provide a robust, flexible framework in which
+new optimizers can be easily developed.  Similarly, the need is also real for a
+flexible framework that allows new objective functions to be easily implemented
+and optimized with a variety of possible optimizers.
 
-{\bf <TODO: elaborate on above, add a few more sentences.>}
+However, the current landscape of optimization frameworks for machine learning
+is not particularly comprehensive.  A variety of tools such as
+Caffe \cite{jia2014caffe},
+% mxnet ??
+TensorFlow \cite{abadi2016tensorflow},
+and
+Keras \cite{chollet2015}
+have optimization frameworks, but they are limited to SGD-type optimizers and
+only are able to optimize deep neural networks or related structures.  Thus
+expressing arbitrary machine learning objective functions can be difficult or in
+some cases not possible.  Other libraries, like
+% Shogun \cite{shogun}
+%and
+scikit-learn \cite{pedregosa2011scikit},
+do have optimizers implementation, but generally not in a coherent framework and
+often the implementations may be specific to an individual machine learning
+algorithm.  At a higher level, many programming languages may have generic
+optimizers, like SciPy \cite{jones2014scipy} and MATLAB, but typically these
+optimizers are not suitable for large-scale machine learning tasks where, e.g.,
+calculating the full gradient of all of the data may not be feasible.
 
-{\bf <TODO: mention related work, or really that we haven't found other
-libraries that have a comprehensive infrastructure like the one we have built,
-and that what we have built is more flexible and generic}
+Given this situation, we have developed a flexible optimization infrastructure
+in the {\bf mlpack} C++ machine learning library \cite{mlpack2013}.  This
+infrastructure makes it easy to combine nearly any type of optimizer with nearly
+any type of objective function, and has allowed us to minimize the effort
+necessary to both implement new optimizers and to implement new machine learning
+algorithms that depend on optimization.  Since the framework is implemented in
+C++ and uses template metaprogramming, we are able to preserve a clean syntax
+while simultaneously allowing compile-time optimizations that can result in a
+generic optimizer that is equivalently fast to an optimizer written specifically
+for one objective function.
+
+In this short paper, we describe mlpack's optimization infrastructure in detail.
+First, we describe the types of optimization problems we would like to solve,
+which then allows us to build a generic and intuitive API for objective
+functions to implement and optimizers to depend upon.  We show a very simple
+example objective function and optimizer, and a system that we have built to
+detect when a user passes an objective function that cannot be optimized by a
+certain optimizer.  Lastly, we show the optimizers we have implemented in this
+framework, and some example usage of mlpack's optimizers.
+
+%In recent times, machine learning conferences have focused more and more on
+%optimization techniques.  For instance, this year at ICML (ICML 2017), every
+%single session had at least one track devoted solely to optimization
+%techniques.
+%<TODO: add some more filler here.  Show that optimization is an important part
+%of machine learning, in whatever ways.>
+
+%\begin{itemize}
+%  \item This sequence of events (or situation) motivates the development of a
+%generic, flexible, and fast optimization framework.
+%  \item We have implemented exactly this in {\bf mlpack}, and it's good.
+%\end{itemize}
+
+%{\bf <TODO: elaborate on above, add a few more sentences.>}
+
+%{\bf <TODO: mention related work, or really that we haven't found other
+%libraries that have a comprehensive infrastructure like the one we have built,
+%and that what we have built is more flexible and generic}
 
@@
 
+% I guess to save some space we should group them.
 \begin{itemize}
-  \item {\bf <TODO: add all optimizers>}
+  \item {\bf SGD variants:} Vanilla SGD, SCD, SGDR, SMORMS3, AdaGrad, AdaDelta, RMSProp, Adam, AdaMax
+  \item {\bf Quasi-Newton variants:} LBFGS, IQN, AugLagrangian
+  \item {\bf Genetic variants:} CNE, CMA-ES
+  \item {\bf Simulated Annealing}: Vanilla SA
   \item {\bf <TODO: add all things that we can optimize>}
@@
 
+\bibliographystyle{plain}
+\bibliography{paper}
+
 \end{document}

### 2017-10-17 paper.tex
Update method/optimizer section.

@@
 \begin{itemize}
-  \item {\bf SGD variants:} Vanilla SGD, SCD, SGDR, SMORMS3, AdaGrad, AdaDelta, RMSProp, Adam, AdaMax
-  \item {\bf Quasi-Newton variants:} LBFGS, IQN, AugLagrangian
-  \item {\bf Genetic variants:} CNE, CMA-ES
-  \item {\bf Simulated Annealing}: Vanilla SA
-  \item {\bf <TODO: add all things that we can optimize>}
+  \item {\bf SGD variants:} Stochastic Gradient Descent (SGD), Stochastic
+      Coordinate Descent (SCD), Parallel Stochastic Gradient Descent (Hogwild!),
+      Stochastic Gradient Descent with Restarts (SGDR), SMORMS3, AdaGrad,
+      AdaDelta, RMSProp, Adam, AdaMax
+
+  \item {\bf Quasi-Newton variants:} Limited-memory BFGS (L-BFGS), incremental
+        Quasi-Newton method (IQN), Augmented Lagrangian Method
+
+  \item {\bf Genetic variants:} Conventional Neuro-evolution (CNE), Covariance
+        Matrix Adaptation Evolution Strategy (CMA-ES)
+
+  \item Simulated Annealing
+
+  \item Frank-Wolfe algorithm
+
+  \item {\bf Objective functions:} Neural Networks, Logistic regression,
+      Matrix completion, Neighborhood Components Analysis, Regularized SVD,
+      Reinforcement learning, Softmax regression, Sparse autoencoders,
+      Sparse SVM
 \end{itemize}
+
+In addition, many methods are currently in development and will be released in
+the future.
 

### 2017-10-18 paper.tex
Add simple objective function and policy design pattern.

@@
 \usepackage{threeparttable}
+\usepackage{listings}
 
@@
 
-\usepackage[]{nips_2017}
+\usepackage[final]{nips_2017}
 
@@
 
+For instance, the code below defines a simple sparse function, where each
+dimension has a parabola with a distinct minimum.
+
+\begin{lstlisting}
+class SparseFunction
+{
+  SparseFunction()
+  {
+    intercepts = arma::vec("20 12 15 100");
+    bi = arma::vec("-4 -2 -3 -8");
+  }
+
+  double Evaluate(const arma::mat& p, const size_t i)
+  {
+    return p[i] * p[i] + bi[i] * p[i] + intercepts[i];
+  }
+
+  void Gradient(const arma::mat& p, const size_t i, arma::sp_mat& g)
+  {
+    g.zeros(arma::size(p));
+    g[i] = 2 * p[i] + bi[i];
+  }
+}
+
+\end{lstlisting}
+
+Another important component of the optimization framework design is flexibility.
+To this end, we use a C++ programming paradigm known as policy-based design \cite{Alexandrescu2001}.
+In short, this means that the behavior is easily controllable by the user, simply by specifying template parameters.
+For example, the SGD class, which is used as basis for other stochastic gradient descent methods, accepts a template parameter UpdateType.
+Thus, if the user wants to update the learning rate update strategy that mlpack does not have support for, they merely need to implement a optimizer class, and then they can use the type SGD<MyCustomLearningRate>.
+Additionally, because templates are being used, there is no additional overhead, as there would be when providing this type of support through inheritance or in other languages such as Python or C
+
 \section{Static checking of functions}
@@
 
-  \item Simulated Annealing
+  \item {\bf Annealing variants:} Simulated Annealing
 
-  \item Frank-Wolfe algorithm
+  \item {\bf Else:} Conditional Gradient Descent (Frank-Wolfe algorithm)
 

### 2017-10-18 paper.tex
Remove final tag.
Add language tag (c++).

@@
 
-\usepackage[final]{nips_2017}
+\usepackage[]{nips_2017}
 
@@
 
+\lstset{language=C++}
 \begin{lstlisting}
@@
 Another important component of the optimization framework design is flexibility.
-To this end, we use a C++ programming paradigm known as policy-based design \cite{Alexandrescu2001}.
-In short, this means that the behavior is easily controllable by the user, simply by specifying template parameters.
-For example, the SGD class, which is used as basis for other stochastic gradient descent methods, accepts a template parameter UpdateType.
-Thus, if the user wants to update the learning rate update strategy that mlpack does not have support for, they merely need to implement a optimizer class, and then they can use the type SGD<MyCustomLearningRate>.
-Additionally, because templates are being used, there is no additional overhead, as there would be when providing this type of support through inheritance or in other languages such as Python or C
+To this end, we use a C++ programming paradigm known as policy-based design
+\cite{Alexandrescu2001}. In short, this means that the behavior is easily
+controllable by the user, simply by specifying template parameters. For example,
+the SGD class, which is used as basis for other stochastic gradient descent
+methods, accepts a template parameter UpdateType. Thus, if the user wants to
+update the learning rate update strategy that mlpack does not have support for,
+they merely need to implement a optimizer class, and then they can use the type
+SGD<MyCustomLearningRate>. Additionally, because templates are being used, there
+is no additional overhead, as there would be when providing this type of support
+through inheritance or in other languages such as Python or C.
 

### 2017-10-18 paper.tex
Fill out section 2; switch to minted for code listings.

@@
 \usepackage{threeparttable}
-\usepackage{listings}
+\usepackage{minted}
 
@@
 
-\section{mlpack: a scalable C++ machine learning library}
+\section{The mlpack machine learning library}
 
-\begin{itemize}
-  \item Some quick background on what mlpack is and what its goals are.
-Probably the list of four goals is good here (though maybe not in itemized
-format).
-  \item Discuss why we've chosen C++, and how policy-based design gives us some
-really nice flexibility.  (Maybe there is not fully space for that.)
-\end{itemize}
+mlpack is a C++ machine learning library with an emphasis on speed, flexibility,
+and ease of use \cite{mlpack2013}; it has been continuously developed since
+2007.  mlpack uses the Armadillo library \cite{sanderson2016armadillo} for
+linear algebra and matrix primitives.  Both mlpack and Armadillo exploit C++
+language features such as policy-based design and template metaprogramming to
+provide compile-time optimizations that result in fast code \cite{pasc2017},
+while retaining a simple and easy-to-use interface \cite{icopust2017}.
+
+Many of the speedups in mlpack depend on the technique of {\it policy-based
+design} \cite{Alexandrescu2001}.  With this design paradigm, mlpack
+provides a number of classes with modular functionality.  As a simple example,
+mlpack's mean shift clustering implementation (as of version 2.2.5) takes three
+template parameters:
+
+\begin{minted}{c++}
+  template<bool UseKernel, typename KernelType, typename MatType>
+  class MeanShift;
+\end{minted}
+
+Thus a user wishing to perform kernelized mean shift clustering with the
+Gaussian kernel might simply use the class
+
+\begin{minted}{c++}
+  MeanShift<true, GaussianKernel, arma::mat>
+\end{minted}
+
+\noindent where {\tt arma::mat} is Armadillo's dense matrix type and {\tt
+GaussianKernel} is a class that provides one simple {\tt Evaluate()} function
+that returns the kernel value between its two arguments.  Since the {\tt
+KernelType} argument is a template argument, any class can be used---it does not
+need to be functionality supplied with mlpack.  Therefore, a user can simply
+implement their own {\tt KernelType} class with the single {\tt Evaluate()}
+function and use that as a template argument to the {\tt MeanShift} class, and
+the compiler will generate a specialized implementation using the custom kernel.
+Similar support could be accomplished by, e.g., function pointers (and this is
+often done in other languages); however, templates allow us to avoid the runtime
+cost of dereferencing the function pointer by generating code that directly
+calls the correct function, which can be a non-negligible cost especially when
+the function is called many times.
+
+mlpack uses policy-based design throughout the library, so that any
+functionality can be easily extended or modified by the user without needing to
+dig into the internals of the code.  Our optimization framework is built around
+this paradigm, allowing for fast prototyping of either new optimizers or new
+objective functions.
 
@@
 
-\lstset{language=C++}
-\begin{lstlisting}
+\begin{minted}{c++}
 class SparseFunction
@@
 }
-
-\end{lstlisting}
+\end{minted}
 

### 2017-10-18 paper.tex
Add two more features.

@@
 \centering
-    \begin{tabular}{@{} cl*{4}c @{}}
-        & & \multicolumn{4}{c}{} \\[2ex]
+    \begin{tabular}{@{} cl*{6}c @{}}
+        & & \multicolumn{6}{c}{} \\[2ex]
         & & \rot{has framework}
           & \rot{\shortstack[l]{problem-independent\\constraints}}
-          & \rot{supports x}
-          & \rot{supports x} \\
-        \cmidrule{2-6}
-        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE  \\
-        & vowpal wabbit      &  - & - & -  & -  \\
-        & TensorFlow         &  - & -  & -  & -  \\
-        & scikit-learn       &  - & -  & -  & - \\
-        & scipy              &  - & \LEFTcircle  & \LEFTcircle  & \LEFTcircle \\
-        \cmidrule[1pt]{2-6}
+          & \rot{batches}
+          & \rot{reusability}
+          & \rot{x}
+          & \rot{y} \\
+        \cmidrule{2-8}
+        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE  \\
+        & vowpal wabbit      & - & - & \LEFTcircle  & \CIRCLE & - & - \\
+        & TensorFlow         & - & -  & \CIRCLE  & \LEFTcircle & - & -  \\
+        & scikit-learn       & - & -  & \CIRCLE  & \LEFTcircle & - & - \\
+        & scipy              & - & \LEFTcircle  & -  & \LEFTcircle & - & - \\
+        \cmidrule[1pt]{2-8}
     \end{tabular}
@@
 and ease of use \cite{mlpack2013}; it has been continuously developed since
-2007.  mlpack uses the Armadillo library \cite{sanderson2016armadillo} for
+2007.  mlpack uses the Armadillo library \cite{pasc2017} for
 linear algebra and matrix primitives.  Both mlpack and Armadillo exploit C++
@@
 \begin{itemize}
-  \item {\tt Evaluate()}
-  \item {\tt Gradient()}
-  \item {\tt PartialGradient()} 
+  \item {\tt Evaluate()} Evaluate a function.
+  \item {\tt Gradient()}: Evaluate the gradient of a function.
+
+  \item {\tt PartialGradient()} Evaluate the partial (sparse) gradient of a
+      function.
+
   \item ...

### 2017-10-19 paper.tex
Step by step guide to write simple optimization function (not sure we have to do it step by step).

@@
 \begin{itemize}
-  \item {\tt Evaluate()} Evaluate a function.
-  \item {\tt Gradient()}: Evaluate the gradient of a function.
+  \item {\tt Evaluate()}: Evaluate a function with the given function parameter.
 
-  \item {\tt PartialGradient()} Evaluate the partial (sparse) gradient of a
-      function.
+  \item {\tt Gradient()}: Evaluate the gradient of a function with the given
+      function parameter and return the gradient.
+
+  \item {\tt PartialGradient()}: Evaluate the partial (sparse) gradient of a
+      function with the given parameter and return the partial gradient.
 
@@
 
-For instance, the code below defines a simple sparse function, where each
-dimension has a parabola with a distinct minimum.
+This section details the usage of mlpack's optimization framework. In this
+example, we would like to minimize a simple sparse function, where each
+dimension has a parabola with a distinct minimum. In this example, we will show
+how we can use mlpack’s framework for this task, and obtain the objective of the
+sparse function shown below.
 
@@
 \end{minted}
+
+First, we would like to define {\tt Optimize()} which takes the function we like
+to optimize and the function parameters as an Armadillo dense matrix type as
+input.
+
+\begin{minted}{c++}
+template<typename FunctionType, typename ParameterType>
+double Optimize(DecomposableFunctionType& function, ParameterType& parameter);
+\end{minted}
+
+Now, we can use {\tt Gradient()} to write a simple stochastic gradient descent
+method with a static learning rate to optimize the function parameter.
+
+\begin{minted}{c++}
+for (size_t i = 0; i < 300; ++i)
+{
+  const size_t functionIndex = i % 4;
+  function.Gradient(parameter, functionIndex, gradient);
+  parameter -= 0.1 * gradient;
+}
+\end{minted}
+
+Lastly, we would like to calculate the final objective. This can be done by
+iterating over the parameter space using {\tt Evaluate()}.
+
+\begin{minted}{c++}
+double objective = 0;
+for (size_t i = 0; i < 4; ++i)
+  objective += function.Evaluate(parameter, i);
+\end{minted}
+
+Using this code, we can compile the code, and when we run it we receive the
+following output:
+
+xxxx
+
+This simple example, of course, does not discuss all the intricacies like a more
+complex learning rate update routine, but instead presents how the optimization
+framework could be used in a simple data science context. Adapting the example
+to a real-life application would be more straightforward.
 
@@
 
-  \item {\bf Else:} Conditional Gradient Descent (Frank-Wolfe algorithm)
+  \item {\bf Other:} Conditional Gradient Descent (Frank-Wolfe algorithm)
 

### 2017-10-19 paper.tex
Extend conclusion

@@
 have optimization frameworks, but they are limited to SGD-type optimizers and
-only are able to optimize deep neural networks or related structures.  Thus
+are only able to optimize deep neural networks or related structures.  Thus
 expressing arbitrary machine learning objective functions can be difficult or in
@@
 and new objective functions to be optimized inside of the {\bf mlpack} machine
-learning library.  This supports a
+learning library.
+
+The framework provided by {\bf mlpack} supports a wide array of special cases,
+and offers specialized algorithms that outperform their classic generic
+alternatives.
 

### 2017-10-19 paper.tex
Refactor table, add a few more parts to it. Tilt the text a bit less for easy reading, use \adjustbox to make the width right.

@@
 
-\newcommand*\rot{\rotatebox{90}}
+\usepackage{adjustbox}
+\usepackage{array}
+
 \newcommand*\OK{\ding{51}}
+
+\newcolumntype{R}[2]{%
+  >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
+  l%
+  <{\egroup}%
+}
+\newcommand*\rot{\multicolumn{1}{R{60}{2em}}}
 
@@
   Symantec Corporation \\
+  \texttt{ryan@ratml.org} \\
 }
@@
 optimizers are not suitable for large-scale machine learning tasks where, e.g.,
-calculating the full gradient of all of the data may not be feasible.
+calculating the full gradient of all of the data may not be feasible.  Table
+\ref{tab:features} provides a rough overview of the capabilities of many common
+machine learning libraries and languages.
 
@@
 \centering
-    \begin{tabular}{@{} cl*{6}c @{}}
-        & & \multicolumn{6}{c}{} \\[2ex]
+    \begin{tabular}{@{} cl*{7}c @{}}
+%  \begin{tabular}{ccccccc}
+        & & \multicolumn{7}{c}{} \\[2ex]
+            % If there is any coherent framework at all, this is true.
         & & \rot{has framework}
-          & \rot{\shortstack[l]{problem-independent\\constraints}}
+            % If there is any support for constrained optimization, this is
+            % true.
+          & \rot{constraints}
+            % If the optimization framework can do mini-batch, this is true.
           & \rot{batches}
-          & \rot{reusability}
-          & \rot{x}
-          & \rot{y} \\
-        \cmidrule{2-8}
-        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE  \\
-        & vowpal wabbit      & - & - & \LEFTcircle  & \CIRCLE & - & - \\
-        & TensorFlow         & - & -  & \CIRCLE  & \LEFTcircle & - & -  \\
-        & scikit-learn       & - & -  & \CIRCLE  & \LEFTcircle & - & - \\
-        & scipy              & - & \LEFTcircle  & -  & \LEFTcircle & - & - \\
-        \cmidrule[1pt]{2-8}
+            % If I can implement any arbitrary function to be optimized, this is
+            % true.
+          & \rot{arbitrary functions}
+            % If I can implement any new optimization technique to use, this is
+            % true.
+          & \rot{arbitrary optimizers}
+            % If the framework could take advantage of when the gradient is
+            % sparse, this is true.
+          & \rot{sparse gradients}
+            % If the framework can handle categorical/discrete variables for
+            % optimization, this is true.
+          & \rot{categorical} \\
+        \cmidrule{2-9}
+  \hline
+        % It might be reasonable to say mlpack categorical support is only
+        % partial, but I am not sure exactly where we draw the line.
+        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE \\
+        % The Shogun toolbox has a fairly nice framework, but it doesn't support
+        % sparse gradients or categorical features.  It also does not appear to
+        % support constraints.
+        & shogun             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
+\\
+        % VW doesn't appear to have any framework whatsoever and the code is
+        % awful, but it does support batches and categorical features.
+        & vowpal wabbit      & - & - & \CIRCLE  & - & - & - &
+\CIRCLE \\
+        % TensorFlow has a few optimizers, but they are all SGD-related.  You
+        % can write most objectives easily (but some very hard), and categorical
+        % support might be possible but would not be easy.
+        & TensorFlow         & \CIRCLE & -  & \CIRCLE  & \LEFTcircle & - &
+\LEFTcircle & -  \\
+        % Caffe has a nice framework, but it's only for SGD-related optimizers.
+        % I think I could write a new one, but it is not the easiest thing in
+        % the world.
+        & caffe             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+& - & - \\
+        % Keras is restricted to neural networks and SGD-like optimizers.  I
+        % don't know that it is possible to easily write a new optimizer.
+        & keras             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+& - & - \\
+        % sklearn has a few optimizer frameworks, but they are all in different
+        % places and have somewhat different support.
+        & scikit-learn       & \LEFTcircle & - & \LEFTcircle  & \LEFTcircle & -
+& - & - \\
+        % scipy has some nice optimizer framework but it does not support
+        % batches or some of the more complex functionality.  And you can't
+        % write your own.
+        & scipy              & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
+        % MATLAB is very similar to scipy.
+        & MATLAB             & \CIRCLE & \CIRCLE & - & \CIRCLE & - & - & - \\
+        \cmidrule[1pt]{2-9}
     \end{tabular}
-    \begin{tablenotes}\footnotesize
-\item[*] Library feature comparison: \CIRCLE=\text{provides feature}, \LEFTcircle=\text{partially provides feature}, -=\text{does not provide feature}
-\end{tablenotes}
+%   \begin{tablenotes}\footnotesize
+\caption{
+Library feature comparison: \CIRCLE = provides feature,
+\LEFTcircle = partially provides feature, - = does not provide feature.
+{\it has framework} indicates that the library has some kind of generic
+optimization framework; {\it constraints} and {\it batches} indicate support for
+batches; {\it arbitrary functions} means arbitrary objective functions are
+easily implemented; {\it arbitrary optimizers} means arbitrary optimizers are
+easily implemented; {\it sparse gradient} indicates whether the framework can
+natively take advantage of sparse gradients; and {\it categorical} refers to
+whether support for categorical features exists.
+}
+\label{tab:features}
+\vspace*{-1.5em}
 \end{table}

### 2017-10-19 paper.tex
Add transition to Section 4.

@@
 function to be optimized meets the required assumptions of the optimizers, and
-encode those requirements.  Since we are using templates and C++, the resulting
+encode those requirements.
+Since we are using templates and C++, the resulting
 code generated by the compiler can be identical to what a developer would write
@@
 
-{\bf <TODO: maybe we should add a table with optimizer types that work with
-certain assumptions.  Maybe this can be a table, maybe a diagram, but some kind
-of visual aid to split the space of objective function types would be nice.>}
+In some cases, some subclasses of functions can still be optimized with more
+general optimizers.  For instance, separable objective functions can still be
+optimized with an optimizer that does not specifically support separable
+functions.  Similarly, a sparse differentiable function may be optimized with
+any optimizer that can handle differentiable functions; however, that optimizer
+may not take advantage of the sparseness.
+
+Now, with an understanding of the types of functions that we wish to support and
+their individual characteristics, we can develop an API that a given function
+should implement the relevant parts of.
 

### 2017-10-19 paper.tex
Function type details

@@
 cost of dereferencing the function pointer by generating code that directly
-calls the correct function, which can be a non-negligible cost especially when
-the function is called many times.
+calls the correct function. This cost can be non-negligible, especially in
+situations where the function is called many times, such as the {\tt Evaluate()}
+function of an optimization problem.
 
@@
 may (or must) implement, depending on the subclass of objective function.
+
+At the time of writing, we offer a few implementations of this API, each
+designed to cover a special case.
+
+\begin{itemize}
+  \item{\tt FunctionType}: Requires the implementation of {\tt Evaluate()} and
+  {\tt Gradient()}. This is a generic function that is used by the
+  {\tt GradientDescent} optimizer.
+  \item{\tt DecomposableFunctionType}: In addition to {\tt FunctionType}'s
+  requirements, requires the implementation of {NumFunctions()}, which returns
+  the number of parts the optimization problem can be decomposed into. The other
+  two functions ({\tt Gradient(), Evaluate()}) accept an index as argument and
+  only evaluate one of the decomposable parts. This is used by our {\tt AdaGrad}
+  implementations, some {\tt SGD} implementations, {\tt RMSProp} and
+  {\tt SMORMS3}.
+  \item{\tt SparseFunctionType}: Similar to {\tt DecomposableFunctionType},
+  however this function type returns a sparse gradient, which makes it ideal for
+  sparse, parallelizable optimization problems. This function type is used by
+  {\bf mlpack}'s parallel implementation of SGD, {\tt ParallelSGD}.
+  \item{\tt LagrangianFunctionType}: This function type fits constrained
+  optimization problems. It requires the implementation of {\tt NumConstraints},
+  {\tt EvaluateConstraint()} and {\tt GradientConstraint}, which are required by
+  optimizers like {\tt AugLagrangian} and {\tt LRSDP}.
+\end{itemize}
+
+In addition to these function types, certain optimizers require specific
+functionality to achieve their goal. The {\tt FunctionType} API offers users the
+flexibility to define a custom function signature. As an example, {\bf mlpack}'s
+implementation of L-BFGS requires a function that exposes a
+{\tt GetInitialPoint} function.
 
@@
 and offers specialized algorithms that outperform their classic generic
-alternatives.
+ (alternatives.
 

### 2017-10-21 paper.tex
Fix missing parentheses
Use two column style for the code snippets.

@@
 \usepackage{minted}
+\usepackage{caption}
+\usepackage{parcolumns}
 
@@
 implementation of L-BFGS requires a function that exposes a
-{\tt GetInitialPoint} function.
+{\tt GetInitialPoint()} function.
 
@@
 
-{\bf <TODO: add very simple example function>}
-
 This section details the usage of mlpack's optimization framework. In this
@@
 
-\begin{minted}{c++}
-class SparseFunction
-{
-  SparseFunction()
-  {
-    intercepts = arma::vec("20 12 15 100");
+\begin{figure}[H]
+\begin{minipage}[t]{0.49\textwidth}
+\centering
+\begin{minted}[fontsize=\footnotesize]{c++}
+class ObjectiveFunction {
+ ObjectiveFunction() {
+  in = arma::vec(
+    "20 12 15 100");
   bi = arma::vec("-4 -2 -3 -8");
@@
 
-  double Evaluate(const arma::mat& p, const size_t i)
-  {
-    return p[i] * p[i] + bi[i] * p[i] + intercepts[i];
+ void Shuffel() {
+  order = arma::shuffle(
+   arma::linspace<arma::uvec>(0, 3, 4));
  }
 
-  void Gradient(const arma::mat& p, const size_t i, arma::sp_mat& g)
-  {
+ size_t NumFunctions() { return 4; }
+
+ double Evaluate(const arma::mat& p,
+  size_t s, size_t b) {
+  double cost = 0;
+  for (size_t i = s; i < s + b; i++) {
+   cost += std::pow(p(i), 2) +
+    bi(order(i)) * p(order(i)) +
+    in(order(i));
+  }
+  return cost;
+ }
+
+ void Gradient(const arma::mat& p,
+  size_t s, size_t b, arma::sp_mat& g) {
   g.zeros(arma::size(p));
-    g[i] = 2 * p[i] + bi[i];
+
+  for (size_t i = s; i < s + b; i++) {
+   g(i) += 2 * p(i) + bi(order(i));
  }
@@
 \end{minted}
+\vspace{2\baselineskip}
+% \captionof{listing}{Distinct parabola sparse function.}
+\end{minipage}
+\begin{minipage}[t]{0.49\textwidth}
+\centering
+\begin{minted}[fontsize=\footnotesize]{c++}
+class ObjectiveFunction {
+ ObjectiveFunction() {
+  in = arma::vec(
+    "20 12 15 100");
+  bi = arma::vec("-4 -2 -3 -8");
+ }
 
-First, we would like to define {\tt Optimize()} which takes the function we like
-to optimize and the function parameters as an Armadillo dense matrix type as
-input.
+ void Shuffel() {
+  order = arma::shuffle(
+   arma::linspace<arma::uvec>(0, 3, 4));
+ }
 
-\begin{minted}{c++}
-template<typename FunctionType, typename ParameterType>
-double Optimize(DecomposableFunctionType& function, ParameterType& parameter);
-\end{minted}
+ size_t NumFunctions() { return 4; }
 
-Now, we can use {\tt Gradient()} to write a simple stochastic gradient descent
-method with a static learning rate to optimize the function parameter.
+ double Evaluate(const arma::mat& p,
+  size_t s, size_t b) {
+  double cost = 0;
+  for (size_t i = s; i < s + b; i++) {
+   cost += std::pow(p(i), 2) +
+    bi(order(i)) * p(order(i)) +
+    in(order(i));
+  }
+  return cost;
+ }
 
-\begin{minted}{c++}
-for (size_t i = 0; i < 300; ++i)
-{
-  const size_t functionIndex = i % 4;
-  function.Gradient(parameter, functionIndex, gradient);
-  parameter -= 0.1 * gradient;
+ void Gradient(const arma::mat& p,
+  size_t s, size_t b, arma::sp_mat& g) {
+  g.zeros(arma::size(p));
+
+  for (size_t i = s; i < s + b; i++) {
+   g(i) += 2 * p(i) + bi(order(i));
+ }
 }
 \end{minted}
-
-Lastly, we would like to calculate the final objective. This can be done by
-iterating over the parameter space using {\tt Evaluate()}.
-
-\begin{minted}{c++}
-double objective = 0;
-for (size_t i = 0; i < 4; ++i)
-  objective += function.Evaluate(parameter, i);
-\end{minted}
+% \captionof{listing}{Optimizer class.}
+\end{minipage}
+\captionof{listing}{not relevant}
+\end{figure}
 

### 2017-10-21 paper.tex
Add simple optimizer class.

@@
 \usepackage{threeparttable}
-\usepackage{minted}
+\usepackage[newfloat]{minted}
 \usepackage{caption}
@@
 This section details the usage of mlpack's optimization framework. In this
-example, we would like to minimize a simple sparse function, where each
-dimension has a parabola with a distinct minimum. In this example, we will show
-how we can use mlpack’s framework for this task, and obtain the objective of the
-sparse function shown below.
+example, we would like to minimize a simple function, where each dimension has a
+parabola with a distinct minimum. In this example, we will show how we can use
+mlpack’s framework for this task, and obtain the objective of the function shown
+below.
 
-\begin{figure}[H]
-\begin{minipage}[t]{0.49\textwidth}
-\centering
+\begin{listing}[H]
 \begin{minted}[fontsize=\footnotesize]{c++}
@@
  ObjectiveFunction() {
-  in = arma::vec(
-    "20 12 15 100");
+  in = arma::vec("20 12 15 100");
   bi = arma::vec("-4 -2 -3 -8");
@@
 
- void Shuffel() {
-  order = arma::shuffle(
-   arma::linspace<arma::uvec>(0, 3, 4));
- }
+ void Shuffel() { order = arma::shuffle(arma::linspace<arma::uvec>(0, 3, 4)); }
 
@@
 
- double Evaluate(const arma::mat& p,
-  size_t s, size_t b) {
+ double Evaluate(const arma::mat& para, size_t s, size_t bs) {
   double cost = 0;
-  for (size_t i = s; i < s + b; i++) {
-   cost += std::pow(p(i), 2) +
-    bi(order(i)) * p(order(i)) +
-    in(order(i));
-  }
+  for (size_t i = s; i < s + bs; i++)
+   cost += std::pow(para(i), 2) + bi(order(i)) * para(order(i)) + in(order(i));
   return cost;
@@
 
- void Gradient(const arma::mat& p,
-  size_t s, size_t b, arma::sp_mat& g) {
+ void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
   g.zeros(arma::size(p));
+  for (size_t i = s; i < s + bs; i++)
+   g(i) += 2 * para(i) + bi(order(i));
+  g /= bs;
+ }
 
-  for (size_t i = s; i < s + b; i++) {
-   g(i) += 2 * p(i) + bi(order(i));
- }
+ arma::vec in, bi; arma::uvec order;
 }
 \end{minted}
-\vspace{2\baselineskip}
-% \captionof{listing}{Distinct parabola sparse function.}
-\end{minipage}
-\begin{minipage}[t]{0.49\textwidth}
-\centering
+\caption{Simple objective function where each dimension has a parabola with a
+  distinct minimum}
+\end{listing}
+\begin{listing}[H]
 \begin{minted}[fontsize=\footnotesize]{c++}
-class ObjectiveFunction {
- ObjectiveFunction() {
-  in = arma::vec(
-    "20 12 15 100");
-  bi = arma::vec("-4 -2 -3 -8");
+class SimpleOptimizer {
+ SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) {}
+
+ template<typename FunctionType>
+ double Optimize(FunctionType& function, arma::mat& parameter) {
+  arma::mat gradient;
+  for (size_t i = 0; i < 300; ++i)
+  {
+   const size_t funIdx = i % 4;
+   function.Gradient(parameter, funIdx, gradient, batchSize);
+   parameter -= lr * gradient;
+
+   function.Shuffel();
   }
 
- void Shuffel() {
-  order = arma::shuffle(
-   arma::linspace<arma::uvec>(0, 3, 4));
+  return function.Evaluate(parameter, 0, function.NumFunctions());
  }
-
- size_t NumFunctions() { return 4; }
-
- double Evaluate(const arma::mat& p,
-  size_t s, size_t b) {
-  double cost = 0;
-  for (size_t i = s; i < s + b; i++) {
-   cost += std::pow(p(i), 2) +
-    bi(order(i)) * p(order(i)) +
-    in(order(i));
-  }
-  return cost;
- }
-
- void Gradient(const arma::mat& p,
-  size_t s, size_t b, arma::sp_mat& g) {
-  g.zeros(arma::size(p));
-
-  for (size_t i = s; i < s + b; i++) {
-   g(i) += 2 * p(i) + bi(order(i));
- }
+ size_t bs;  double lr;
 }
 \end{minted}
-% \captionof{listing}{Optimizer class.}
-\end{minipage}
-\captionof{listing}{not relevant}
-\end{figure}
+\caption{Simple stochastic gradient descent method with a static number of
+iterations to optimize the function parameter ({\tt bs} defines the batch size
+and {\tt lr} the learning rate). )}
+\end{listing}
 

### 2017-10-21 paper.tex
Add simple optimizer class.

@@
 
-\begin{minted}{c++}
+\begin{minted}[fontsize=\footnotesize]{c++}
   template<bool UseKernel, typename KernelType, typename MatType>
@@
 
-\begin{minted}{c++}
+\begin{minted}[fontsize=\footnotesize]{c++}
   MeanShift<true, GaussianKernel, arma::mat>
@@
 
+% \vspace*{-0.3em}
 \begin{listing}[H]
@@
 \end{minted}
+\vspace{-0.6em}
 \caption{Simple objective function where each dimension has a parabola with a
   distinct minimum}
+\vspace{-0.8em}
 \end{listing}
+\vspace{-0.8em}
 \begin{listing}[H]
@@
 class SimpleOptimizer {
- SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) {}
+ SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) { }
 
@@
   arma::mat gradient;
-  for (size_t i = 0; i < 300; ++i)
-  {
+  for (size_t i = 0; i < 300; ++i) {
    const size_t funIdx = i % 4;
@@
    parameter -= lr * gradient;
-
    function.Shuffel();
   }
-
   return function.Evaluate(parameter, 0, function.NumFunctions());
@@
 \end{minted}
+\vspace{-0.6em}
 \caption{Simple stochastic gradient descent method with a static number of
@@
 and {\tt lr} the learning rate). )}
+\vspace{-0.6em}
 \end{listing}
+\vspace{-0.8em}
+Using this code, we can compile the code, and when we run it
 
-Using this code, we can compile the code, and when we run it we receive the
-following output:
+\begin{minted}[fontsize=\footnotesize]{c++}
+ObjectiveFunction function; arma::mat parmeter("0 0 0 0;")
+SimpleOptimizer optimizer;
+std::cout << optimizer.Optimize(function, parameter);
+\end{minted}
+we receive the following output:
 

### 2017-10-21 paper.tex
Refactor paper to make it 7 pages, finish unfinished sections.
The basic tricks I used to get the size down:

 * Add \vspace*{-0.5em} (or whatever value) to remove vertical space---but not
   so much the paper is cramped.
 * Inline code in a few places to avoid extra line breaks.
 * Use \footnotesize for code listings, which hopefully aren't too small.
 * Put code directly into paper instead of into a listing that LaTeX can't split
   across pages.
 * Find different words or shorter phrases to remove the last line of paragraphs
   where the line was short.
 * Tilt column headers in Table 1 to

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2017}
+%\usepackage[final]{nips_2017}
 
@@
 }
-\newcommand*\rot{\multicolumn{1}{R{60}{2em}}}
+\newcommand*\rot{\multicolumn{1}{R{30}{2em}}}
 
@@
   Symantec Corporation \\
+  Atlanta, GA, USA 30318 \\
   \texttt{ryan@ratml.org} \\
@@
 arbitrary optimization problems that may arise all throughout machine learning
-problems.  You care because:
-
-\begin{itemize}
-  \item It's easy to implement new optimizers.
-  \item It's easy to implement new functions to be optimized.
-  \item This system is more comprehensive than other techniques that may focus
-only on neural networks or single problems.
-  \item This system is also more specific and offers more room for optimization
-than super general systems like scipy.optimize().
-\end{itemize}
-
-This abstract will need to be rewritten.
+problems.  In this paper we present the generic optimization framework that we
+have designed for mlpack.  A key priority in the design was ease of
+implementation of both new optimizers and new objective functions to be
+optimized; therefore, implementation of a new optimizer requires only one
+method and implementation of a new objective function requires at most four
+functions.  This leads to simple and intuitive code, and for fast prototyping
+and experimentation, this is of paramount importance.  When compared to
+optimization frameworks of other libraries, we find that mlpack's supports more
+types of objective functions, is able to make optimizations that other
+frameworks do not, and seamlessly supports user-defined objective functions
+{\it and} optimizers.
+%\begin{itemize}
+%  \item It's easy to implement new optimizers.
+%  \item It's easy to implement new functions to be optimized.
+%  \item This system is more comprehensive than other techniques that may focus
+%only on neural networks or single problems.
+%  \item This system is also more specific and offers more room for optimization
+%than super general systems like scipy.optimize().
+%\end{itemize}
+%
+%This abstract will need to be rewritten.
 \end{abstract}
@@
 scikit-learn \cite{pedregosa2011scikit},
-do have optimizers implementation, but generally not in a coherent framework and
+do have optimizers, but generally not in a coherent framework and
 often the implementations may be specific to an individual machine learning
@@
 calculating the full gradient of all of the data may not be feasible.  Table
-\ref{tab:features} provides a rough overview of the capabilities of many common
-machine learning libraries and languages.
+\ref{tab:features} provides a rough overview.
 
@@
 algorithms that depend on optimization.  Since the framework is implemented in
-C++ and uses template metaprogramming, we are able to preserve a clean syntax
+C++ and uses template metaprogramming, we are able to preserve clean syntax
 while simultaneously allowing compile-time optimizations that can result in a
-generic optimizer that is equivalently fast to an optimizer written specifically
-for one objective function.
+generic optimizer that is as fast as an optimizer written for one specific objective function.
 
@@
 \label{tab:features}
-\vspace*{-1.5em}
+\vspace*{-1.8em}
 \end{table}
 
+\vspace*{-0.3em}
 \section{The mlpack machine learning library}
+\vspace*{-0.2em}
 
@@
 
+\vspace*{-0.4em}
 \begin{minted}[fontsize=\footnotesize]{c++}
@@
 \end{minted}
+\vspace*{-0.4em}
 
@@
 Gaussian kernel might simply use the class
+{\footnotesize \mintinline{c++}{MeanShift<true, GaussianKernel, arma::mat>}}
+\noindent where {\tt arma::mat} is Armadillo's dense matrix type and {\tt
+GaussianKernel} is a class that provides one simple {\tt Evaluate()} method.
 
-\begin{minted}[fontsize=\footnotesize]{c++}
-  MeanShift<true, GaussianKernel, arma::mat>
-\end{minted}
-
-\noindent where {\tt arma::mat} is Armadillo's dense matrix type and {\tt
-GaussianKernel} is a class that provides one simple {\tt Evaluate()} function
-that returns the kernel value between its two arguments.  Since the {\tt
-KernelType} argument is a template argument, any class can be used---it does not
-need to be functionality supplied with mlpack.  Therefore, a user can simply
+Since the {\tt KernelType} argument is a template argument, any class can be
+used---it does not need to be part of mlpack.  Therefore, a user can simply
 implement their own {\tt KernelType} class with the single {\tt Evaluate()}
-function and use that as a template argument to the {\tt MeanShift} class, and
-the compiler will generate a specialized implementation using the custom kernel.
-Similar support could be accomplished by, e.g., function pointers (and this is
-often done in other languages); however, templates allow us to avoid the runtime
-cost of dereferencing the function pointer by generating code that directly
-calls the correct function. This cost can be non-negligible, especially in
-situations where the function is called many times, such as the {\tt Evaluate()}
-function of an optimization problem.
+method and use that as a template argument to the {\tt MeanShift} class, and
+the compiler will generate a specialized implementation using the custom
+kernel.  Similar support could be accomplished by, e.g., function pointers (and
+this is often done in other languages); however, templates allow us to avoid
+the runtime cost of dereferencing the function pointer by generating code that
+directly calls the correct method. This cost can be non-negligible, especially
+in situations where the method is called many times, such as the {\tt
+Evaluate()} method of an optimization problem.
 
@@
 
+\vspace*{-0.3em}
 \section{Requirements for optimizers and functions}
+\vspace*{-0.2em}
 
@@
 
+\vspace*{-0.5em}
 \begin{equation}
@@
 \end{equation}
+\vspace*{-1.3em}
 
-\noindent for any function $f(x)$ that takes some vector input $x$.  However, it
+\noindent for any function $f(x)$ that takes some vector input $x$.  But it
 is impossible to implement something both so generic and fast---for instance,
@@
 approaches (in general), so we must design an abstraction that is able to
-simultaneously generalize to many problem types, as well as able to take
-advantage of accelerations and optimizations.
+simultaneously generalize to many problem types, as well as take advantage of
+accelerations and optimizations.
 
-Let us split the class of functions to be optimized into some subclasses:
+Let us describe the class of functions to be optimized with some non-exclusive
+properties:
 
-\begin{itemize}
+\vspace*{-0.4em}
+\begin{itemize} \itemsep -1pt
   \item {\bf arbitrary}: no assumptions can be made on $f(x)$
@@
 values
-  \item {\bf numeric}: all elements of $x$ take values in $\mathcal{R}$
+  %\item {\bf numeric}: all elements of $x$ take values in $\mathcal{R}$
   \item {\bf sparse}: the gradient $f'(x)$ or $f_i(x)$ (for a separable
@@
 \end{itemize}
+\vspace*{-0.4em}
 
@@
 In some cases, some subclasses of functions can still be optimized with more
-general optimizers.  For instance, separable objective functions can still be
-optimized with an optimizer that does not specifically support separable
-functions.  Similarly, a sparse differentiable function may be optimized with
-any optimizer that can handle differentiable functions; however, that optimizer
-may not take advantage of the sparseness.
+general optimizers.  For example, separable functions can still be optimized
+with an optimizer that does not specifically support them.  Similarly, a sparse
+differentiable function may be optimized with any optimizer that supports
+differentiable functions; however, the sparseness might not be taken advantage
+of.
 
@@
 their individual characteristics, we can develop an API that a given function
-should implement the relevant parts of.
+can implement the relevant parts of.
 
+\vspace*{-0.3em}
 \section{{\tt FunctionType} API}
+\vspace*{-0.2em}
 
 In order to facilitate consistent implementations, we have defined a {\tt
-FunctionType} API that describes all the functions that an objective function
-may (or must) implement, depending on the subclass of objective function.
+FunctionType} API that describes all the methods that an objective function
+may implement.  mlpack offers a few variations of this API, each designed to
+cover some of the function characteristics of the previous section.
+%At the time of writing, we offer a few implementations of this API, each
+%designed to cover a special case.
+%\begin{itemize}
+%  \item{\tt FunctionType}: Requires the implementation of {\tt Evaluate()} and
+%  {\tt Gradient()}. This is a generic function that is used by the
+%  {\tt GradientDescent} optimizer.
+%  \item{\tt DecomposableFunctionType}: In addition to {\tt FunctionType}'s
+%  requirements, requires the implementation of {NumFunctions()}, which returns
+%  the number of parts the optimization problem can be decomposed into. The other
+%  two functions ({\tt Gradient(), Evaluate()}) accept an index as argument and
+%  only evaluate one of the decomposable parts. This is used by our {\tt AdaGrad}
+%  implementations, some {\tt SGD} implementations, {\tt RMSProp} and
+%  {\tt SMORMS3}.
+%  \item{\tt SparseFunctionType}: Similar to {\tt DecomposableFunctionType},
+%  however this function type returns a sparse gradient, which makes it ideal for
+%  sparse, parallelizable optimization problems. This function type is used by
+%  {\bf mlpack}'s parallel implementation of SGD, {\tt ParallelSGD}.
+% We'll kind of sweep constrained optimization under the rug a little bit, there
+% are a lot of extra complexities we don't have space for.
+%  \item{\tt LagrangianFunctionType}: This function type fits constrained
+%  optimization problems. It requires the implementation of {\tt NumConstraints},
+%  {\tt EvaluateConstraint()} and {\tt GradientConstraint}, which are required by
+%  optimizers like {\tt AugLagrangian} and {\tt LRSDP}.
+%\end{itemize}
+Any {\tt FunctionType} to be optimized requires the implementation of an {\tt
+Evaluate()} method.  The interface used for that can be one of the following
+two methods:
 
-At the time of writing, we offer a few implementations of this API, each
-designed to cover a special case.
+\vspace*{-0.3em}
+\begin{minted}[fontsize=\footnotesize]{c++}
+// For non-separable objectives.
+double Evaluate(const arma::mat& parameters);
+// For separable objectives.
+double Evaluate(const arma::mat& parameters,
+                const size_t start,
+                const size_t batchSize);
+\end{minted}
+\vspace*{-0.3em}
 
-\begin{itemize}
-  \item{\tt FunctionType}: Requires the implementation of {\tt Evaluate()} and
-  {\tt Gradient()}. This is a generic function that is used by the
-  {\tt GradientDescent} optimizer.
-  \item{\tt DecomposableFunctionType}: In addition to {\tt FunctionType}'s
-  requirements, requires the implementation of {NumFunctions()}, which returns
-  the number of parts the optimization problem can be decomposed into. The other
-  two functions ({\tt Gradient(), Evaluate()}) accept an index as argument and
-  only evaluate one of the decomposable parts. This is used by our {\tt AdaGrad}
-  implementations, some {\tt SGD} implementations, {\tt RMSProp} and
-  {\tt SMORMS3}.
-  \item{\tt SparseFunctionType}: Similar to {\tt DecomposableFunctionType},
-  however this function type returns a sparse gradient, which makes it ideal for
-  sparse, parallelizable optimization problems. This function type is used by
-  {\bf mlpack}'s parallel implementation of SGD, {\tt ParallelSGD}.
-  \item{\tt LagrangianFunctionType}: This function type fits constrained
-  optimization problems. It requires the implementation of {\tt NumConstraints},
-  {\tt EvaluateConstraint()} and {\tt GradientConstraint}, which are required by
-  optimizers like {\tt AugLagrangian} and {\tt LRSDP}.
-\end{itemize}
+Both of these methods should calculate the objective given the parameters
+matrix {\tt parameters}; however, the second overload is for {\it separable}
+functions, and should calculate the partial objective starting at the separable
+function indexed by {\tt start} and calculate {\tt batchSize} partial objectives
+and return the sum.  Functions implementing the first overload are used by
+optimizers like the {\tt GradientDescent} optimizer; functions implementing the
+second are used by SGD-like optimizers.
 
-In addition to these function types, certain optimizers require specific
-functionality to achieve their goal. The {\tt FunctionType} API offers users the
-flexibility to define a custom function signature. As an example, {\bf mlpack}'s
-implementation of L-BFGS requires a function that exposes a
-{\tt GetInitialPoint()} function.
+Note that, importantly, it is easy to calculate the objective for a
+non-separable function with the second overload just by setting {\tt start} to
+$0$ and {\tt batchSize} to $1$.  Therefore, it is easy to make an `adapter' that
+can allow separable optimizers to work with non-separable functions (and vice
+versa).
 
-{\bf <TODO: we can add a list of each function and what it does.  A basic list
-is below.  We can elaborate on this significantly.>}
+Next, any {\it differentiable} function must implement some {\tt Gradient()}
+method.
 
-\begin{itemize}
-  \item {\tt Evaluate()}: Evaluate a function with the given function parameter.
+\vspace*{-0.3em}
+\begin{minted}[fontsize=\footnotesize]{c++}
+// For non-separable differentiable sparse and non-sparse functions.
+template<typename GradType>
+void Gradient(const arma::mat& parameters, GradType& gradient);
+// For separable differentiable sparse and non-sparse functions.
+template<typename GradType>
+void Gradient(const arma::mat& parameters,
+              const size_t start,
+              GradType& gradient,
+              const size_t batchSize);
+\end{minted}
+\vspace*{-0.3em}
 
-  \item {\tt Gradient()}: Evaluate the gradient of a function with the given
-      function parameter and return the gradient.
+Both of these methods should calculate the gradient and place the results into
+the matrix object {\tt gradient}.  Note that the method accepts a template
+parameter {\tt GradType}, which may be {\tt arma::mat} (dense Armadillo matrix)
+or {\tt arma::sp\_mat} (sparse Armadillo matrix).  This allows support for both
+sparse-supporting and non-sparse-supporting optimizers.\footnote{One could write
+a non-templated {\tt Gradient()} method for just {\tt arma::mat}, and it would
+work fine for non-sparse optimizers.  But to us it seems just as easy to
+templatize it.}
 
-  \item {\tt PartialGradient()}: Evaluate the partial (sparse) gradient of a
-      function with the given parameter and return the partial gradient.
+Next, if the objective function is {\it partially differentiable}, we can
+implement the following method:
 
-  \item ...
-  \item constraints?
-  \item discrete variable support?
-  \item batch support
-\end{itemize}
+\vspace*{-0.3em}
+\begin{minted}[fontsize=\footnotesize]{c++}
+// For partially differentiable sparse and non-sparse functions.
+template<typename GradType>
+void PartialGradient(const arma::mat& parameters,
+                     const size_t j,
+                     GradType& gradient);
+\end{minted}
+\vspace*{-0.3em}
 
-A user may implement as many (or as few) of these functions as they would like,
-and this will allow them to use a larger set of optimizers.
+This should calculate the gradient of the parameters {\tt parameters} with
+respect to the parameter {\tt j} and store the results (either sparse or dense)
+in the {\tt gradient} matrix object.
+
+In addition, {\it separable} and {\it partially differentiable} functions must
+implement the {\tt NumFunctions()} and {\tt NumFeatures()} functions,
+respectively:
+
+\vspace*{-0.3em}
+\begin{minted}[fontsize=\footnotesize]{c++}
+// For separable functions: return the number of parts the optimization problem
+// can be decomposed into.
+size_t NumFunctions();
+// For partially differentiable functions: return the number of partial derivatives.
+size_t NumFeatures();
+\end{minted}
+\vspace*{-0.3em}
+
+Finally, {\it separable} functions must implement the method
+{\footnotesize \mintinline{c++}{void Shuffle()}}, which shuffles the ordering
+of the functions (note that the optimizer is not required to call it).  This is
+useful for data-based problems, where it may not be desirable to loop over the
+separable objective functions---which usually correspond to individual data
+points---in the same order.
+
+These simple functions, however, do not specify how to handle {\it categorical}
+or {\it bounded} functions.  In those cases, the {\tt FunctionType} should
+accept, in its constructor, its bounds and which dimensions are categorical or
+numeric.  Since constraint types can differ greatly, it is up to an individual
+optimizer to define the format in which it should receive its constraints.  More
+information can be found in the documentation of mlpack's {\tt LRSDP}, {\tt
+AugLagrangian}, and {\tt FrankWolfe} optimizers.
+
+%% The constraints are part of the optimizer themselves.
+% EvaluateConstraint()
+% NumConstraints()
+
+%\begin{itemize}
+%  \item {\tt double Evaluate(const arma::mat& parameters)} {\it (arbitrary
+%functions)}: return the objective function given the parameters.
+%  \item {\tt void Gradient(const arma::mat& parameters, arma::mat& gradient)}
+%{\it (differentiable)}: compute the gradient and store it in {\tt gradient}
+%given the parameters.
+%  \item {\tt double Evaluate(const arma::mat& parameters, const size\_t start,
+%const size\_t batchSize)} {\it (separable)}: compute a partial objective
+%function, starting with separable function {\tt start} and computing {\tt batchSize}
+%functions.
+%  \item {\tt void Gradient(arma::mat& parameters, const size\_t start,
+%arma::mat& gradient, const size\_t batchSize)} {\it (separable)}: compute a part
+%of the gradient starting with separable function {\tt start} and computing {\tt
+%batchSize} individual gradients, storing the result in {\tt gradient}.
+
+\vspace*{-0.3em}
+\section{Optimizer API}
+\vspace*{-0.4em}
+
+In addition to implementing functions, users can also add new optimizers easily
+if they implement an optimizer with a simple API.  Fortunately, the requirements
+for implementing optimizers are much simpler than for objective functions.  An
+optimizer must implement only the method
+
+\vspace*{-0.5em}
+\begin{minted}[fontsize=\footnotesize]{c++}
+template<typename FunctionType>
+double Optimize(FunctionType& function,
+                arma::mat& parameters);
+\end{minted}
+\vspace*{-0.5em}
+
+The {\tt Optimize()} method should check that the given {\tt FunctionType}
+satisfies the assumptions the optimizer makes (see Section \ref{sec:static}) and
+optimize the given function {\tt function}, storing the best set of parameters
+in the matrix {\tt parameters} and returning the best objective value.
+
+\vspace*{-0.3em}
+\section{Example function and optimizer}
+\vspace*{-0.4em}
 
@@
 % \vspace*{-0.3em}
-\begin{listing}[H]
+%\begin{listing}[H]
 \begin{minted}[fontsize=\footnotesize]{c++}
@@
 
- void Shuffel() { order = arma::shuffle(arma::linspace<arma::uvec>(0, 3, 4)); }
-
   size_t NumFunctions() { return 4; }
+  void Shuffle() { order = arma::shuffle(arma::uvec("0 1 2 3 ")); }
 
@@
     for (size_t i = s; i < s + bs; i++)
-   cost += std::pow(para(i), 2) + bi(order(i)) * para(order(i)) + in(order(i));
+      cost += (para(i) * para(i)) + bi(order(i)) * para(order(i)) + in(order(i));
     return cost;
@@
   void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
-  g.zeros(arma::size(p));
+    g.zeros(para.n_rows, para.n_cols);
     for (size_t i = s; i < s + bs; i++)
-   g(i) += 2 * para(i) + bi(order(i));
-  g /= bs;
+      g(i) += (1.0 / bs) * 2 * para(i) + bi(order(i));
   }
@@
   arma::vec in, bi; arma::uvec order;
-}
+};
 \end{minted}
-\vspace{-0.6em}
-\caption{Simple objective function where each dimension has a parabola with a
-  distinct minimum}
-\vspace{-0.8em}
-\end{listing}
-\vspace{-0.8em}
-\begin{listing}[H]
+%\vspace{-0.6em}
+%\caption{Simple objective function where each dimension has a parabola with a
+%  distinct minimum}
+%\vspace{-0.8em}
+%\end{listing}
+%\vspace{-0.8em}
+%\begin{listing}[H]
 \begin{minted}[fontsize=\footnotesize]{c++}
@@
     arma::mat gradient;
-  for (size_t i = 0; i < 300; ++i) {
+    for (size_t i = 0; i < 300; i += bs) {
       const size_t funIdx = i % 4;
@@
       parameter -= lr * gradient;
-   function.Shuffel();
+      if (funIdx == 0) { function.Shuffle(); }
     }
@@
   }
- size_t bs;  double lr;
-}
+  size_t bs; double lr; // Store batch size and learning rate internally.
+};
 \end{minted}
-\vspace{-0.6em}
-\caption{Simple stochastic gradient descent method with a static number of
-iterations to optimize the function parameter ({\tt bs} defines the batch size
-and {\tt lr} the learning rate). )}
-\vspace{-0.6em}
-\end{listing}
-\vspace{-0.8em}
+%\vspace{-0.6em}
+%\caption{Simple stochastic gradient descent method with a static number of
+%iterations to optimize the function parameter ({\tt bs} defines the batch size
+%and {\tt lr} the learning rate). )}
+%\vspace{-0.6em}
+%\end{listing}
+%\vspace{-0.8em}
 Using this code, we can compile the code, and when we run it
 
+\vspace*{-0.4em}
 \begin{minted}[fontsize=\footnotesize]{c++}
-ObjectiveFunction function; arma::mat parmeter("0 0 0 0;")
+  ObjectiveFunction function; arma::mat parameter("0 0 0 0;")
   SimpleOptimizer optimizer;
@@
 \end{minted}
+\vspace*{-0.4em}
+
 we receive the following output:
 
+\vspace*{-0.4em}
+{\footnotesize
+\begin{verbatim}
 xxxx
+\end{verbatim}
+}
+\vspace*{-0.4em}
 
@@
 
-\section{Static checking of functions}
-
-{\bf <TODO: probably a better section title>}
+\section{Statically checking function properties}
+\label{sec:static}
 
@@
 support to determine what methods a type has available.  Similarly, we can also
-use static compile-time constants to denote the methods that are require by a
+use static compile-time constants to denote the methods that are required by a
 specific optimizer.  This is implemented via SFINAE \cite{sfinae}.  A {\tt
 static\_assert()} is raised when a given objective function does not have the
-required methods for a given optimizer.  {\bf <TODO: implement this, so we are
-not lying.  But it should be easy enough, maybe an hour of work max.>}
+required methods for a given optimizer.
 
 For instance, when attempting to use the L-BFGS optimizer without having a
-{\tt Gradient()} function implemented, the user will receive the following error
-message:
+{\tt Gradient()} function implemented, the user will receive a (comparatively)
+simple error message of the form:
 
-{\bf <TODO: put error message here.  If we are short on space, let's cut this
-example and make this section shorter.>}
+\vspace*{-0.4em}
+{\footnotesize
+\begin{verbatim}
+  error: static assertion failed: the FunctionType does not have a correct
+      definition of a Gradient() function
+\end{verbatim}
+}
+\vspace*{-0.4em}
 
@@
 
+\vspace*{-0.4em}
 % I guess to save some space we should group them.
-\begin{itemize}
+\begin{itemize} \itemsep -1pt
   \item {\bf SGD variants:} Stochastic Gradient Descent (SGD), Stochastic
@@
 
-  \item {\bf Other:} Conditional Gradient Descent (Frank-Wolfe algorithm)
+  \item {\bf Other:} Conditional Gradient Descent, Frank-Wolfe algorithm
 
@@
 observation to provide an easy framework for both implementing new optimizers
-and new objective functions to be optimized inside of the {\bf mlpack} machine
+and new objective functions to be optimized inside of the mlpack machine
 learning library.
 
-The framework provided by {\bf mlpack} supports a wide array of special cases,
-and offers specialized algorithms that outperform their classic generic
- (alternatives.
+The framework provided by mlpack supports a wide array of special cases, and
+already has implemented specialized algorithms that outperform their classic
+generic alternatives.
 

### 2017-10-21 paper.tex
Fix example code issues, add output and SFINAE reference.

@@
 \begin{minted}[fontsize=\footnotesize]{c++}
-class ObjectiveFunction {
+struct ObjectiveFunction {
   ObjectiveFunction() {
@@
     for (size_t i = s; i < s + bs; i++)
-      cost += (para(i) * para(i)) + bi(order(i)) * para(order(i)) + in(order(i));
+      cost += (para(i) * para(i)) + bi(order(i)) * para(i) + in(order(i));
     return cost;
@@
 
-  void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
+  void Gradient(const arma::mat& para, size_t s, arma::mat& g, size_t bs) {
     g.zeros(para.n_rows, para.n_cols);
@@
 \begin{minted}[fontsize=\footnotesize]{c++}
-class SimpleOptimizer {
-  SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) { }
+struct SimpleOptimizer {
+  SimpleOptimizer(size_t bs = 1, double lr = 0.02) : bs(bs), lr(lr) { }
 
@@
     arma::mat gradient;
-    for (size_t i = 0; i < 300; i += bs) {
+    for (size_t i = 0; i < 5000; i += bs) {
       const size_t funIdx = i % 4;
-      function.Gradient(parameter, funIdx, gradient, batchSize);
+      if (funIdx == 0) { function.Shuffle(); }
+      function.Gradient(parameter, funIdx, gradient, bs);
       parameter -= lr * gradient;
-      if (funIdx == 0) { function.Shuffle(); }
     }
@@
 \begin{minted}[fontsize=\footnotesize]{c++}
-  ObjectiveFunction function; arma::mat parameter("0 0 0 0;")
+  ObjectiveFunction function; arma::mat parameter("0 0 0 0;");
   SimpleOptimizer optimizer;
-  std::cout << optimizer.Optimize(function, parameter);
+  std::cout << "objective: " << optimizer.Optimize(function, parameter);
 \end{minted}
@@
 
-we receive the following output:
+we receive the output: {\tt objective: 123.8 }
 
-\vspace*{-0.4em}
-{\footnotesize
-\begin{verbatim}
-xxxx
-\end{verbatim}
-}
+The final value of the objective function should be close to the optimal value,
+which is the sum of values at the vertices of the parabolas.
 \vspace*{-0.4em}
@@
 use static compile-time constants to denote the methods that are required by a
-specific optimizer.  This is implemented via SFINAE \cite{sfinae}.  A {\tt
-static\_assert()} is raised when a given objective function does not have the
-required methods for a given optimizer.
+specific optimizer.  This is implemented via SFINAE
+\cite{Vandevoorde2002sfinae}.  A {\tt static\_assert()} is raised when a given
+objective function does not have the required methods for a given optimizer.
 

### 2017-10-21 paper.tex
Minor grammer fix.
Add a bit about the optimizer, fix to fit in 6 pages.

@@
 arbitrary optimization problems that may arise all throughout machine learning
-problems.  In this paper we present the generic optimization framework that we
+problems.  In this paper, we present the generic optimization framework that we
 have designed for mlpack.  A key priority in the design was ease of
@@
 template<typename FunctionType>
-double Optimize(FunctionType& function,
-                arma::mat& parameters);
+double Optimize(FunctionType& function, arma::mat& parameters);
 \end{minted}
@@
 
-\vspace*{-0.3em}
+\vspace*{-0.5em}
 \section{Example function and optimizer}
-\vspace*{-0.4em}
+\vspace*{-0.6em}
 
@@
 example, we would like to minimize a simple function, where each dimension has a
-parabola with a distinct minimum. In this example, we will show how we can use
-mlpack’s framework for this task, and obtain the objective of the function shown
-below.
+parabola with a distinct minimum. In this example, we show how to use mlpack’s
+framework for this task, and minimize the function below.
 
@@
 %\begin{listing}[H]
+\vspace*{-0.3em}
 \begin{minted}[fontsize=\footnotesize]{c++}
 struct ObjectiveFunction {
-  ObjectiveFunction() {
+  ObjectiveFunction() { // A separable function consisting of four quadratics.
     in = arma::vec("20 12 15 100");
@@
   size_t NumFunctions() { return 4; }
-  void Shuffle() { order = arma::shuffle(arma::uvec("0 1 2 3 ")); }
+  void Shuffle() { ord = arma::shuffle(arma::uvec("0 1 2 3")); }
 
@@
     for (size_t i = s; i < s + bs; i++)
-      cost += (para(i) * para(i)) + bi(order(i)) * para(i) + in(order(i));
+      cost += (para(ord[i]) * para(ord[i]) + bi(ord[i]) * para[i] + in(ord[i]);
     return cost;
@@
     for (size_t i = s; i < s + bs; i++)
-      g(i) += (1.0 / bs) * 2 * para(i) + bi(order(i));
+      g(ord[i]) += (1.0 / bs) * 2 * para(ord[i]) + bi(ord[i]);
   }
 
-  arma::vec in, bi; arma::uvec order;
+  arma::vec in /* intercepts */, bi /* coeffs */; arma::uvec ord /* function order */;
 };
 \end{minted}
-%\vspace{-0.6em}
+\vspace*{-0.3em}
+
+Note that in this code, we maintain an ordering with the vector {\tt order}; in
+other situations, such as training neural networks, we could simply shuffle the columns of the data
+matrix in {\tt Shuffle()}.  This objective function will work with any mlpack
+optimizer that supports separable or differentiable functions; this includes
+all SGD-like optimizers, L-BFGS, simulated annealing, and others.
+
+Next, we wish to define a simple example optimizer that can be used with {\tt
+ObjectiveFunction} and other mlpack objective functions.  For this, we must
+implement only an {\tt Optimize()} method, and a constructor to set some
+parameters.  The code is given below.
+
+\vspace{-0.3em}
 %\caption{Simple objective function where each dimension has a parabola with a
@@
     for (size_t i = 0; i < 5000; i += bs) {
-      const size_t funIdx = i % 4;
-      if (funIdx == 0) { function.Shuffle(); }
-      function.Gradient(parameter, funIdx, gradient, bs);
+      if ((i % function.NumFunctions() == 0) { function.Shuffle(); }
+      function.Gradient(parameter, i % function.NumFunctions(), gradient, bs);
       parameter -= lr * gradient;
@@
 \end{minted}
-%\vspace{-0.6em}
+\vspace{-0.3em}
 %\caption{Simple stochastic gradient descent method with a static number of
@@
 %\vspace{-0.8em}
-Using this code, we can compile the code, and when we run it
+
+Note that for the sake of brevity we have omitted checks on the batch size
+(this optimizer assumes that {\tt function.NumFunctions()} is a multiple of the
+batch size) and other typical parts of real implementations, as well as any
+static type checking to ensure separability and differentiability.  Still, {\tt
+SimpleOptimizer} can work with any mlpack objective function satisfying those
+conditions.  This includes mlpack's neural network code, logistic regression,
+and other objective functions.
+
+Now, we can find a minimum of {\tt ObjectiveFunction} with {\tt
+SimpleOptimizer} using this code:
 
@@
 
-we receive the output: {\tt objective: 123.8 }
+When we run this code, we receive the output: {\tt objective: 123.8 }
 
@@
 which is the sum of values at the vertices of the parabolas.
-\vspace*{-0.4em}
-
 This simple example, of course, does not discuss all the intricacies like a more
@@
 framework could be used in a simple data science context. Adapting the example
-to a real-life application would be more straightforward.
+to a real-life application would be straightforward.
 
-Another important component of the optimization framework design is flexibility.
-To this end, we use a C++ programming paradigm known as policy-based design
-\cite{Alexandrescu2001}. In short, this means that the behavior is easily
-controllable by the user, simply by specifying template parameters. For example,
-the SGD class, which is used as basis for other stochastic gradient descent
-methods, accepts a template parameter UpdateType. Thus, if the user wants to
-update the learning rate update strategy that mlpack does not have support for,
-they merely need to implement a optimizer class, and then they can use the type
-SGD<MyCustomLearningRate>. Additionally, because templates are being used, there
-is no additional overhead, as there would be when providing this type of support
-through inheritance or in other languages such as Python or C.
+An important point to re-emphasize is that the use of templates and
+policy-based design allows easy control of behavior by users, simply by
+specifying template parameters---or writing custom classes when needed.  In
+addition, there is no runtime performance penalty for this flexibility, as
+there would be when providing this type of support through inheritance or in
+other languages such as Python or C.
 
+%Another important component of the optimization framework design is flexibility.
+%To this end, we use a C++ programming paradigm known as policy-based design
+%\cite{Alexandrescu2001}. In short, this means that the behavior is easily
+%controllable by the user, simply by specifying template parameters. For example,
+%the SGD class, which is used as basis for other stochastic gradient descent
+%methods, accepts a template parameter UpdateType. Thus, if the user wants to
+%update the learning rate update strategy that mlpack does not have support for,
+%they merely need to implement a optimizer class, and then they can use the type
+%SGD<MyCustomLearningRate>. Additionally, because templates are being used, there
+%is no additional overhead, as there would be when providing this type of support
+%through inheritance or in other languages such as Python or C.
+
+\vspace*{-0.4em}
 \section{Statically checking function properties}
 \label{sec:static}
+\vspace*{-0.3em}
 
@@
 }
-\vspace*{-0.4em}
+\vspace*{-0.9em}
 
 \section{Supported optimizers and functions in mlpack}
+\vspace*{-0.3em}
 
@@
 
+\vspace*{-0.4em}
 \section{Conclusion}
+\vspace*{-0.3em}
 
@@
 learning library.
-
 The framework provided by mlpack supports a wide array of special cases, and

### 2017-10-21 paper.tex
Fix clamps and update optimizer output.

@@
     for (size_t i = s; i < s + bs; i++)
-      cost += (para(ord[i]) * para(ord[i]) + bi(ord[i]) * para[i] + in(ord[i]);
+      cost += para(ord[i]) * para(ord[i]) + bi(ord[i]) * para(ord[i]) + in(ord[i]);
     return cost;
@@
     for (size_t i = 0; i < 5000; i += bs) {
-      if ((i % function.NumFunctions() == 0) { function.Shuffle(); }
+      if (i % function.NumFunctions() == 0) { function.Shuffle(); }
       function.Gradient(parameter, i % function.NumFunctions(), gradient, bs);
@@
 
-When we run this code, we receive the output: {\tt objective: 123.8 }
+When we run this code, we receive the output: {\tt objective: 123.75 }
 

### 2017-10-21 paper.tex
Add references for each library in Table 1.

@@
 algorithm.  At a higher level, many programming languages may have generic
-optimizers, like SciPy \cite{jones2014scipy} and MATLAB, but typically these
+optimizers, like SciPy \cite{jones2014scipy} and MATLAB \cite{mathworks2017OTB}, but typically these
 optimizers are not suitable for large-scale machine learning tasks where, e.g.,
@@
         % partial, but I am not sure exactly where we draw the line.
-        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE \\
+        & mlpack \cite{mlpack2013}            & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE \\
         % The Shogun toolbox has a fairly nice framework, but it doesn't support
@@
         % support constraints.
-        & shogun             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
+        & shogun \cite{sonnenburg2010shogun}             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
 \\
@@
         % awful, but it does support batches and categorical features.
-        & vowpal wabbit      & - & - & \CIRCLE  & - & - & - &
+        & vowpal wabbit \cite{Langford2007VW}      & - & - & \CIRCLE  & - & - & - &
 \CIRCLE \\
@@
         % support might be possible but would not be easy.
-        & TensorFlow         & \CIRCLE & -  & \CIRCLE  & \LEFTcircle & - &
+        & TensorFlow \cite{abadi2016tensorflow}        & \CIRCLE & -  & \CIRCLE  & \LEFTcircle & - &
 \LEFTcircle & -  \\
@@
         % the world.
-        & caffe             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+        & caffe \cite{jia2014caffe}           & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
 & - & - \\
@@
         % don't know that it is possible to easily write a new optimizer.
-        & keras             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+        & keras \cite{chollet2015}            & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
 & - & - \\
@@
         % places and have somewhat different support.
-        & scikit-learn       & \LEFTcircle & - & \LEFTcircle  & \LEFTcircle & -
+        & scikit-learn \cite{pedregosa2011scikit}       & \LEFTcircle & - & \LEFTcircle  & \LEFTcircle & -
 & - & - \\
@@
         % write your own.
-        & scipy              & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
+        & scipy \cite{jones2014scipy}             & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
         % MATLAB is very similar to scipy.
-        & MATLAB             & \CIRCLE & \CIRCLE & - & \CIRCLE & - & - & - \\
+        & MATLAB \cite{mathworks2017OTB}            & \CIRCLE & \CIRCLE & - & \CIRCLE & - & - & - \\
         \cmidrule[1pt]{2-9}

### 2017-10-22 paper.tex
Change phrasing in table legend

--- 2017-10-21 paper.tex
@@
 arbitrary optimization problems that may arise all throughout machine learning
-problems.  In this paper, we present the generic optimization framework that we
+problems.  In this paper we present the generic optimization framework that we
 have designed for mlpack.  A key priority in the design was ease of
@@
 \caption{
-Library feature comparison: \CIRCLE = provides feature,
+Feature comparison: \CIRCLE = provides feature,
 \LEFTcircle = partially provides feature, - = does not provide feature.
-{\it has framework} indicates that the library has some kind of generic
+{\it has framework} the library has some kind of generic
 optimization framework; {\it constraints} and {\it batches} indicate support for
-batches; {\it arbitrary functions} means arbitrary objective functions are
-easily implemented; {\it arbitrary optimizers} means arbitrary optimizers are
-easily implemented; {\it sparse gradient} indicates whether the framework can
-natively take advantage of sparse gradients; and {\it categorical} refers to
-whether support for categorical features exists.
+constrained problems and batches; {\it arbitrary functions} means arbitrary
+objective functions are easily implemented; {\it arbitrary optimizers} means
+arbitrary optimizers are easily implemented; {\it sparse gradient} indicates
+that the framework can natively take advantage of sparse gradients; and
+{\it categorical} refers to whether support for categorical features exists.
 }
@@
 template<typename FunctionType>
-double Optimize(FunctionType& function, arma::mat& parameters);
+double Optimize(FunctionType& function,
+                arma::mat& parameters);
 \end{minted}
@@
 
-\vspace*{-0.5em}
+\vspace*{-0.3em}
 \section{Example function and optimizer}
-\vspace*{-0.6em}
+\vspace*{-0.4em}
 
@@
 example, we would like to minimize a simple function, where each dimension has a
-parabola with a distinct minimum. In this example, we show how to use mlpack’s
-framework for this task, and minimize the function below.
+parabola with a distinct minimum. In this example, we will show how we can use
+mlpack’s framework for this task, and obtain the objective of the function shown
+below.
 
@@
 %\begin{listing}[H]
-\vspace*{-0.3em}
 \begin{minted}[fontsize=\footnotesize]{c++}
-struct ObjectiveFunction {
-  ObjectiveFunction() { // A separable function consisting of four quadratics.
+class ObjectiveFunction {
+  ObjectiveFunction() {
     in = arma::vec("20 12 15 100");
@@
   size_t NumFunctions() { return 4; }
-  void Shuffle() { ord = arma::shuffle(arma::uvec("0 1 2 3")); }
+  void Shuffle() { order = arma::shuffle(arma::uvec("0 1 2 3 ")); }
 
@@
     for (size_t i = s; i < s + bs; i++)
-      cost += (para(ord[i]) * para(ord[i]) + bi(ord[i]) * para[i] + in(ord[i]);
+      cost += (para(i) * para(i)) + bi(order(i)) * para(order(i)) + in(order(i));
     return cost;
@@
 
-  void Gradient(const arma::mat& para, size_t s, arma::mat& g, size_t bs) {
+  void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
     g.zeros(para.n_rows, para.n_cols);
     for (size_t i = s; i < s + bs; i++)
-      g(ord[i]) += (1.0 / bs) * 2 * para(ord[i]) + bi(ord[i]);
+      g(i) += (1.0 / bs) * 2 * para(i) + bi(order(i));
   }
 
-  arma::vec in /* intercepts */, bi /* coeffs */; arma::uvec ord /* function order */;
+  arma::vec in, bi; arma::uvec order;
 };
 \end{minted}
-\vspace*{-0.3em}
-
-Note that in this code, we maintain an ordering with the vector {\tt order}; in
-other situations, such as training neural networks, we could simply shuffle the columns of the data
-matrix in {\tt Shuffle()}.  This objective function will work with any mlpack
-optimizer that supports separable or differentiable functions; this includes
-all SGD-like optimizers, L-BFGS, simulated annealing, and others.
-
-Next, we wish to define a simple example optimizer that can be used with {\tt
-ObjectiveFunction} and other mlpack objective functions.  For this, we must
-implement only an {\tt Optimize()} method, and a constructor to set some
-parameters.  The code is given below.
-
-\vspace{-0.3em}
+%\vspace{-0.6em}
 %\caption{Simple objective function where each dimension has a parabola with a
@@
 \begin{minted}[fontsize=\footnotesize]{c++}
-struct SimpleOptimizer {
-  SimpleOptimizer(size_t bs = 1, double lr = 0.02) : bs(bs), lr(lr) { }
+class SimpleOptimizer {
+  SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) { }
 
@@
     arma::mat gradient;
-    for (size_t i = 0; i < 5000; i += bs) {
-      if ((i % function.NumFunctions() == 0) { function.Shuffle(); }
-      function.Gradient(parameter, i % function.NumFunctions(), gradient, bs);
+    for (size_t i = 0; i < 300; i += bs) {
+      const size_t funIdx = i % 4;
+      function.Gradient(parameter, funIdx, gradient, batchSize);
       parameter -= lr * gradient;
+      if (funIdx == 0) { function.Shuffle(); }
     }
@@
 \end{minted}
-\vspace{-0.3em}
+%\vspace{-0.6em}
 %\caption{Simple stochastic gradient descent method with a static number of
@@
 %\vspace{-0.8em}
-
-Note that for the sake of brevity we have omitted checks on the batch size
-(this optimizer assumes that {\tt function.NumFunctions()} is a multiple of the
-batch size) and other typical parts of real implementations, as well as any
-static type checking to ensure separability and differentiability.  Still, {\tt
-SimpleOptimizer} can work with any mlpack objective function satisfying those
-conditions.  This includes mlpack's neural network code, logistic regression,
-and other objective functions.
-
-Now, we can find a minimum of {\tt ObjectiveFunction} with {\tt
-SimpleOptimizer} using this code:
+Using this code, we can compile the code, and when we run it
 
@@
 \begin{minted}[fontsize=\footnotesize]{c++}
-  ObjectiveFunction function; arma::mat parameter("0 0 0 0;");
+  ObjectiveFunction function; arma::mat parameter("0 0 0 0;")
   SimpleOptimizer optimizer;
-  std::cout << "objective: " << optimizer.Optimize(function, parameter);
+  std::cout << optimizer.Optimize(function, parameter);
 \end{minted}
@@
 
-When we run this code, we receive the output: {\tt objective: 123.8 }
+we receive the following output:
 
-The final value of the objective function should be close to the optimal value,
-which is the sum of values at the vertices of the parabolas.
+\vspace*{-0.4em}
+{\footnotesize
+\begin{verbatim}
+xxxx
+\end{verbatim}
+}
+\vspace*{-0.4em}
+
 This simple example, of course, does not discuss all the intricacies like a more
@@
 framework could be used in a simple data science context. Adapting the example
-to a real-life application would be straightforward.
+to a real-life application would be more straightforward.
 
-An important point to re-emphasize is that the use of templates and
-policy-based design allows easy control of behavior by users, simply by
-specifying template parameters---or writing custom classes when needed.  In
-addition, there is no runtime performance penalty for this flexibility, as
-there would be when providing this type of support through inheritance or in
-other languages such as Python or C.
+Another important component of the optimization framework design is flexibility.
+To this end, we use a C++ programming paradigm known as policy-based design
+\cite{Alexandrescu2001}. In short, this means that the behavior is easily
+controllable by the user, simply by specifying template parameters. For example,
+the SGD class, which is used as basis for other stochastic gradient descent
+methods, accepts a template parameter UpdateType. Thus, if the user wants to
+update the learning rate update strategy that mlpack does not have support for,
+they merely need to implement a optimizer class, and then they can use the type
+SGD<MyCustomLearningRate>. Additionally, because templates are being used, there
+is no additional overhead, as there would be when providing this type of support
+through inheritance or in other languages such as Python or C.
 
-%Another important component of the optimization framework design is flexibility.
-%To this end, we use a C++ programming paradigm known as policy-based design
-%\cite{Alexandrescu2001}. In short, this means that the behavior is easily
-%controllable by the user, simply by specifying template parameters. For example,
-%the SGD class, which is used as basis for other stochastic gradient descent
-%methods, accepts a template parameter UpdateType. Thus, if the user wants to
-%update the learning rate update strategy that mlpack does not have support for,
-%they merely need to implement a optimizer class, and then they can use the type
-%SGD<MyCustomLearningRate>. Additionally, because templates are being used, there
-%is no additional overhead, as there would be when providing this type of support
-%through inheritance or in other languages such as Python or C.
-
-\vspace*{-0.4em}
 \section{Statically checking function properties}
 \label{sec:static}
-\vspace*{-0.3em}
 
@@
 use static compile-time constants to denote the methods that are required by a
-specific optimizer.  This is implemented via SFINAE
-\cite{Vandevoorde2002sfinae}.  A {\tt static\_assert()} is raised when a given
-objective function does not have the required methods for a given optimizer.
+specific optimizer.  This is implemented via SFINAE \cite{sfinae}.  A {\tt
+static\_assert()} is raised when a given objective function does not have the
+required methods for a given optimizer.
 
@@
 }
-\vspace*{-0.9em}
+\vspace*{-0.4em}
 
 \section{Supported optimizers and functions in mlpack}
-\vspace*{-0.3em}
 
@@
 
-\vspace*{-0.4em}
 \section{Conclusion}
-\vspace*{-0.3em}
 
@@
 learning library.
+
 The framework provided by mlpack supports a wide array of special cases, and

### 2017-10-22 paper.tex
Add affiliation and email address
Capitalize library names. Change expressions.

@@
   Shikhar Bhardwaj \\
+  Delhi Technological University \\
+  Delhi - 110042 \\
+  \texttt{shikhar_bt2k15@dtu.ac.in} \\
   \AND
@@
 method and implementation of a new objective function requires at most four
-functions.  This leads to simple and intuitive code, and for fast prototyping
-and experimentation, this is of paramount importance.  When compared to
+functions.  This leads to simple and intuitive code, which, for fast prototyping
+and experimentation, is of paramount importance.  When compared to
 optimization frameworks of other libraries, we find that mlpack's supports more
@@
 C++ and uses template metaprogramming, we are able to preserve clean syntax
-while simultaneously allowing compile-time optimizations that can result in a
-generic optimizer that is as fast as an optimizer written for one specific objective function.
+while simultaneously allowing compile-time speedups that can result in a
+generic optimizer that is as fast as an optimizer written for one specific
+objective function.
 
@@
         % support constraints.
-        & shogun \cite{sonnenburg2010shogun}             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
+        & Shogun \cite{sonnenburg2010shogun}             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
 \\
@@
         % awful, but it does support batches and categorical features.
-        & vowpal wabbit \cite{Langford2007VW}      & - & - & \CIRCLE  & - & - & - &
+        & Vowpal Wabbit \cite{Langford2007VW}      & - & - & \CIRCLE  & - & - & - &
 \CIRCLE \\
@@
         % the world.
-        & caffe \cite{jia2014caffe}           & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+        & Caffe \cite{jia2014caffe}           & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
 & - & - \\
@@
         % don't know that it is possible to easily write a new optimizer.
-        & keras \cite{chollet2015}            & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
+        & Keras \cite{chollet2015}            & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
 & - & - \\
@@
         % write your own.
-        & scipy \cite{jones2014scipy}             & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
+        & SciPy \cite{jones2014scipy}             & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
         % MATLAB is very similar to scipy.
@@
 Both of these methods should calculate the gradient and place the results into
-the matrix object {\tt gradient}.  Note that the method accepts a template
+the matrix object {\tt gradient} that is passed as an argument.
+Note that the method accepts a template
 parameter {\tt GradType}, which may be {\tt arma::mat} (dense Armadillo matrix)

### 2017-10-22 paper.tex
Slight change in phrasing, Section7

@@
 debug a problem when they implement an objective function or gradient.  To
-provide good error message output, we can use C++'s template metaprogramming
+ improve error message output, we can use C++'s template metaprogramming
 support to determine what methods a type has available.  Similarly, we can also
@@
 \cite{Vandevoorde2002sfinae}.  A {\tt static\_assert()} is raised when a given
-objective function does not have the required methods for a given optimizer.
+objective function does not implement the methods required by the optimizer used.
 

### 2017-11-17 paper.tex
Escape mail address for the submission.

@@
 \author{
+  Ryan R. Curtin \\
+  Symantec Corporation \\
+  Atlanta, GA, USA 30318 \\
+  \texttt{ryan@ratml.org} \\
+  \And
   Shikhar Bhardwaj \\
@@
   Delhi - 110042 \\
-  \texttt{shikhar_bt2k15@dtu.ac.in} \\
-  \AND
-  Yannis Mentekidis \\
-  \AND
+  \texttt{shikhar\char`_bt2k15@dtu.ac.in}
+  \And
   Marcus Edel \\
@@
   \texttt{marcus.edel@fu-berlin.de} \\
-  \AND
-  Ryan R. Curtin \\
-  Symantec Corporation \\
-  Atlanta, GA, USA 30318 \\
-  \texttt{ryan@ratml.org} \\
+  \And
+  Yannis Mentekidis \\
+  mentekid@gmail.com
 }
@@
 
+\vspace*{-0.5cm}
 \begin{abstract}
@@
 
+\vspace*{-0.5cm}
 \section{Introduction}
@@
 template<typename GradType>
-void PartialGradient(const arma::mat& parameters,
-                     const size_t j,
-                     GradType& gradient);
+void PartialGradient(const arma::mat& parameters, const size_t j, GradType& gradient);
 \end{minted}
@@
 
-\vspace*{-0.3em}
+\vspace*{-0.4em}
 \section{Optimizer API}
-\vspace*{-0.4em}
+\vspace*{-0.5em}
 
@@
   ObjectiveFunction() { // A separable function consisting of four quadratics.
-    in = arma::vec("20 12 15 100");
-    bi = arma::vec("-4 -2 -3 -8");
+    in = arma::vec("20 12 15 100"); bi = arma::vec("-4 -2 -3 -8");
   }
@@
 
-\vspace*{-0.4em}
+\vspace*{-0.5em}
 {\footnotesize
@@
 
-  \item {\bf Annealing variants:} Simulated Annealing
-
-  \item {\bf Other:} Conditional Gradient Descent, Frank-Wolfe algorithm
+  \item {\bf Other:} Conditional Gradient Descent, Frank-Wolfe algorithm, Simulated Annealing
 

### 2017-11-17 paper-arxiv.tex
Version with inlined minted and minor word changes for arXiv.

--- 2017-11-17 paper.tex
@@
 \usepackage{threeparttable}
-\usepackage[newfloat]{minted}
+\usepackage[newfloat,draft]{minted}
 \usepackage{caption}
 \usepackage{parcolumns}
+
+\makeatletter
+\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
+    \let\PYG@ul=\relax \let\PYG@tc=\relax%
+    \let\PYG@bc=\relax \let\PYG@ff=\relax}
+\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
+\def\PYG@toks#1+{\ifx\relax#1\empty\else%
+    \PYG@tok{#1}\expandafter\PYG@toks\fi}
+\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
+    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
+\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}
+
+\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
+\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
+\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@vm\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@cs\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
+\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
+\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PYG@tok@nl\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@nd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@ne\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@dl\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@ch\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@fm\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PYG@tok@mb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@cpf\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
+\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@sa\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+
+\def\PYGZbs{\char`\\}
+\def\PYGZus{\char`\_}
+\def\PYGZob{\char`\{}
+\def\PYGZcb{\char`\}}
+\def\PYGZca{\char`\^}
+\def\PYGZam{\char`\&}
+\def\PYGZlt{\char`\<}
+\def\PYGZgt{\char`\>}
+\def\PYGZsh{\char`\#}
+\def\PYGZpc{\char`\%}
+\def\PYGZdl{\char`\$}
+\def\PYGZhy{\char`\-}
+\def\PYGZsq{\char`\'}
+\def\PYGZdq{\char`\"}
+\def\PYGZti{\char`\~}
+% for compatibility with earlier versions
+\def\PYGZat{@}
+\def\PYGZlb{[}
+\def\PYGZrb{]}
+\makeatother
+
+\makeatletter
+\def\PYGdefault@reset{\let\PYGdefault@it=\relax \let\PYGdefault@bf=\relax%
+    \let\PYGdefault@ul=\relax \let\PYGdefault@tc=\relax%
+    \let\PYGdefault@bc=\relax \let\PYGdefault@ff=\relax}
+\def\PYGdefault@tok#1{\csname PYGdefault@tok@#1\endcsname}
+\def\PYGdefault@toks#1+{\ifx\relax#1\empty\else%
+    \PYGdefault@tok{#1}\expandafter\PYGdefault@toks\fi}
+\def\PYGdefault@do#1{\PYGdefault@bc{\PYGdefault@tc{\PYGdefault@ul{%
+    \PYGdefault@it{\PYGdefault@bf{\PYGdefault@ff{#1}}}}}}}
+\def\PYGdefault#1#2{\PYGdefault@reset\PYGdefault@toks#1+\relax+\PYGdefault@do{#2}}
+
+\expandafter\def\csname PYGdefault@tok@gd\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gu\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gt\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gs\endcsname{\let\PYGdefault@bf=\textbf}
+\expandafter\def\csname PYGdefault@tok@gr\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@cm\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@vg\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@vi\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@vm\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@mh\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@cs\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ge\endcsname{\let\PYGdefault@it=\textit}
+\expandafter\def\csname PYGdefault@tok@vc\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@il\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@go\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
+\expandafter\def\csname PYGdefault@tok@cp\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gi\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gh\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ni\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nl\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nn\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@no\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@na\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nb\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nc\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nd\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ne\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nf\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@si\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PYGdefault@tok@s2\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nt\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@nv\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@s1\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@dl\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ch\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@m\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@gp\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sh\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ow\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sx\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@bp\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@c1\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@fm\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@o\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kc\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@c\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@mf\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@err\endcsname{\def\PYGdefault@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PYGdefault@tok@mb\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@ss\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sr\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PYGdefault@tok@mo\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kd\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@mi\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kn\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@cpf\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kr\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@s\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kp\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@w\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PYGdefault@tok@kt\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sc\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sb\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sa\endcsname{\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@k\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PYGdefault@tok@se\endcsname{\let\PYGdefault@bf=\textbf\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PYGdefault@tok@sd\endcsname{\let\PYGdefault@it=\textit\def\PYGdefault@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+
+\def\PYGdefaultZbs{\char`\\}
+\def\PYGdefaultZus{\char`\_}
+\def\PYGdefaultZob{\char`\{}
+\def\PYGdefaultZcb{\char`\}}
+\def\PYGdefaultZca{\char`\^}
+\def\PYGdefaultZam{\char`\&}
+\def\PYGdefaultZlt{\char`\<}
+\def\PYGdefaultZgt{\char`\>}
+\def\PYGdefaultZsh{\char`\#}
+\def\PYGdefaultZpc{\char`\%}
+\def\PYGdefaultZdl{\char`\$}
+\def\PYGdefaultZhy{\char`\-}
+\def\PYGdefaultZsq{\char`\'}
+\def\PYGdefaultZdq{\char`\"}
+\def\PYGdefaultZti{\char`\~}
+% for compatibility with earlier versions
+\def\PYGdefaultZat{@}
+\def\PYGdefaultZlb{[}
+\def\PYGdefaultZrb{]}
+\makeatother
 
@@
 
-\usepackage[final]{nips_2017}
+\usepackage[final,nonatbib]{nips_2017}
 
@@
 \begin{abstract}
+\vspace*{-0.3em}
 The development of the {\bf mlpack} C++ machine learning library
@@
 
-\vspace*{-0.5cm}
+\vspace*{-0.7cm}
 \section{Introduction}
+\vspace*{-0.3em}
 
@@
 example objective function and optimizer, and a system that we have built to
-detect when a user passes an objective function that cannot be optimized by a
-certain optimizer.  Lastly, we show the optimizers we have implemented in this
-framework, and some example usage of mlpack's optimizers.
+detect when an objective function that cannot be optimized by a certain
+optimizer is used.  Lastly, we show the optimizers we have implemented in this
+framework, and example usage of mlpack's optimizers.
 
@@
 that the framework can natively take advantage of sparse gradients; and
-{\it categorical} refers to whether support for categorical features exists.
+{\it categorical} refers to if categorical feature support exists.
 }
@@
 \vspace*{-0.4em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-  template<bool UseKernel, typename KernelType, typename MatType>
-  class MeanShift;
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+  \PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k+kt}{bool} \PYG{n}{UseKernel}\PYG{p}{,} \PYG{k}{typename} \PYG{n}{KernelType}\PYG{p}{,} \PYG{k}{typename} \PYG{n}{MatType}\PYG{o}{\PYGZgt{}}
+  \PYG{k}{class} \PYG{n+nc}{MeanShift}\PYG{p}{;}
+\end{Verbatim}
+}
 \vspace*{-0.4em}
@@
 Gaussian kernel might simply use the class
-{\footnotesize \mintinline{c++}{MeanShift<true, GaussianKernel, arma::mat>}}
+{\footnotesize {\tt MeanShift<true, GaussianKernel, arma::mat>}}
+%\begin{Verbatim}[commandchars=\\\{\}]
+%\PYG{n}{MeanShift}\PYG{o}{\PYGZlt{}}\PYG{n+nb}{true}\PYG{p}{,} \PYG{n}{GaussianKernel}\PYG{p}{,} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZgt{}}
+%\end{Verbatim}
+%}
 \noindent where {\tt arma::mat} is Armadillo's dense matrix type and {\tt
-GaussianKernel} is a class that provides one simple {\tt Evaluate()} method.
+GaussianKernel} provides one simple {\tt Evaluate()} method.
 
@@
 code generated by the compiler can be identical to what a developer would write
-if they were writing an optimizer specifically for the function $f(x)$ at
-hand---this can provide significant efficiency gains.
+if they were writing an optimizer specifically for the function $f(x)$---this
+can give significant efficiency gains.
 
@@
 \vspace*{-0.3em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-// For non-separable objectives.
-double Evaluate(const arma::mat& parameters);
-// For separable objectives.
-double Evaluate(const arma::mat& parameters,
-                const size_t start,
-                const size_t batchSize);
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{c+c1}{// For non\PYGZhy{}separable objectives.}
+\PYG{k+kt}{double} \PYG{n+nf}{Evaluate}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{);}
+\PYG{c+c1}{// For separable objectives.}
+\PYG{k+kt}{double} \PYG{n+nf}{Evaluate}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{,}
+                \PYG{k}{const} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{start}\PYG{p}{,}
+                \PYG{k}{const} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{batchSize}\PYG{p}{);}
+\end{Verbatim}
+}
 \vspace*{-0.3em}
@@
 \vspace*{-0.3em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-// For non-separable differentiable sparse and non-sparse functions.
-template<typename GradType>
-void Gradient(const arma::mat& parameters, GradType& gradient);
-// For separable differentiable sparse and non-sparse functions.
-template<typename GradType>
-void Gradient(const arma::mat& parameters,
-              const size_t start,
-              GradType& gradient,
-              const size_t batchSize);
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{c+c1}{// For non\PYGZhy{}separable differentiable sparse and non\PYGZhy{}sparse functions.}
+\PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k}{typename} \PYG{n}{GradType}\PYG{o}{\PYGZgt{}}
+\PYG{k+kt}{void} \PYG{n}{Gradient}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{,} \PYG{n}{GradType}\PYG{o}{\PYGZam{}} \PYG{n}{gradient}\PYG{p}{);}
+\PYG{c+c1}{// For separable differentiable sparse and non\PYGZhy{}sparse functions.}
+\PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k}{typename} \PYG{n}{GradType}\PYG{o}{\PYGZgt{}}
+\PYG{k+kt}{void} \PYG{n}{Gradient}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{,}
+              \PYG{k}{const} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{start}\PYG{p}{,}
+              \PYG{n}{GradType}\PYG{o}{\PYGZam{}} \PYG{n}{gradient}\PYG{p}{,}
+              \PYG{k}{const} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{batchSize}\PYG{p}{);}
+\end{Verbatim}
+}
 \vspace*{-0.3em}
@@
 \vspace*{-0.3em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-// For partially differentiable sparse and non-sparse functions.
-template<typename GradType>
-void PartialGradient(const arma::mat& parameters, const size_t j, GradType& gradient);
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{c+c1}{// For partially differentiable sparse and non\PYGZhy{}sparse functions.}
+\PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k}{typename} \PYG{n}{GradType}\PYG{o}{\PYGZgt{}}
+\PYG{k+kt}{void} \PYG{n}{PartialGradient}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{,} \PYG{k}{const} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{j}\PYG{p}{,} \PYG{n}{GradType}\PYG{o}{\PYGZam{}} \PYG{n}{gradient}\PYG{p}{);}
+\end{Verbatim}
+}
 \vspace*{-0.3em}
@@
 \vspace*{-0.3em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-// For separable functions: return the number of parts the optimization problem
-// can be decomposed into.
-size_t NumFunctions();
-// For partially differentiable functions: return the number of partial derivatives.
-size_t NumFeatures();
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{c+c1}{// For separable functions: return the number of parts the optimization problem}
+\PYG{c+c1}{// can be decomposed into.}
+\PYG{k+kt}{size\PYGZus{}t} \PYG{n+nf}{NumFunctions}\PYG{p}{();}
+\PYG{c+c1}{// For partially differentiable functions: return the number of partial derivatives.}
+\PYG{k+kt}{size\PYGZus{}t} \PYG{n+nf}{NumFeatures}\PYG{p}{();}
+\end{Verbatim}
+}
 \vspace*{-0.3em}
@@
 Finally, {\it separable} functions must implement the method
-{\footnotesize \mintinline{c++}{void Shuffle()}}, which shuffles the ordering
+{\footnotesize {\tt void Shuffle()}},
+%\begin{Verbatim}[commandchars=\\\{\}]
+%\PYG{k+kt}{void} \PYG{n}{Shuffle}\PYG{p}{()}
+%\end{Verbatim}
+%},
+which shuffles the ordering
 of the functions (note that the optimizer is not required to call it).  This is
@@
 \vspace*{-0.5em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-template<typename FunctionType>
-double Optimize(FunctionType& function, arma::mat& parameters);
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k}{typename} \PYG{n}{FunctionType}\PYG{o}{\PYGZgt{}}
+\PYG{k+kt}{double} \PYG{n}{Optimize}\PYG{p}{(}\PYG{n}{FunctionType}\PYG{o}{\PYGZam{}} \PYG{n}{function}\PYG{p}{,} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameters}\PYG{p}{);}
+\end{Verbatim}
+}
 \vspace*{-0.5em}
@@
 
-\vspace*{-0.5em}
+\vspace*{-0.6em}
 \section{Example function and optimizer}
-\vspace*{-0.6em}
+\vspace*{-0.7em}
 
@@
 %\begin{listing}[H]
-\vspace*{-0.3em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-struct ObjectiveFunction {
-  ObjectiveFunction() { // A separable function consisting of four quadratics.
-    in = arma::vec("20 12 15 100"); bi = arma::vec("-4 -2 -3 -8");
+\vspace*{-0.5em}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{k}{struct} \PYG{n}{ObjectiveFunction} \PYG{p}{\PYGZob{}}
+  \PYG{n}{ObjectiveFunction}\PYG{p}{()} \PYG{p}{\PYGZob{}} \PYG{c+c1}{// A separable function consisting of four quadratics.}
+    \PYG{n}{in} \PYG{o}{=} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{vec}\PYG{p}{(}\PYG{l+s}{\PYGZdq{}20 12 15 100\PYGZdq{}}\PYG{p}{);} \PYG{n}{bi} \PYG{o}{=} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{vec}\PYG{p}{(}\PYG{l+s}{\PYGZdq{}\PYGZhy{}4 \PYGZhy{}2 \PYGZhy{}3 \PYGZhy{}8\PYGZdq{}}\PYG{p}{);}
+  \PYG{p}{\PYGZcb{}}
+
+  \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{NumFunctions}\PYG{p}{()} \PYG{p}{\PYGZob{}} \PYG{k}{return} \PYG{l+m+mi}{4}\PYG{p}{;} \PYG{p}{\PYGZcb{}}
+  \PYG{k+kt}{void} \PYG{n}{Shuffle}\PYG{p}{()} \PYG{p}{\PYGZob{}} \PYG{n}{ord} \PYG{o}{=} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{shuffle}\PYG{p}{(}\PYG{n}{arma}\PYG{o}{::}\PYG{n}{uvec}\PYG{p}{(}\PYG{l+s}{\PYGZdq{}0 1 2 3\PYGZdq{}}\PYG{p}{));} \PYG{p}{\PYGZcb{}}
+
+  \PYG{k+kt}{double} \PYG{n}{Evaluate}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{para}\PYG{p}{,} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{s}\PYG{p}{,} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{bs}\PYG{p}{)} \PYG{p}{\PYGZob{}}
+    \PYG{k+kt}{double} \PYG{n}{cost} \PYG{o}{=} \PYG{l+m+mi}{0}\PYG{p}{;}
+    \PYG{k}{for} \PYG{p}{(}\PYG{k+kt}{size\PYGZus{}t} \PYG{n}{i} \PYG{o}{=} \PYG{n}{s}\PYG{p}{;} \PYG{n}{i} \PYG{o}{\PYGZlt{}} \PYG{n}{s} \PYG{o}{+} \PYG{n}{bs}\PYG{p}{;} \PYG{n}{i}\PYG{o}{++}\PYG{p}{)}
+      \PYG{n}{cost} \PYG{o}{+=} \PYG{n}{para}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{*} \PYG{n}{para}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{+} \PYG{n}{bi}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{*} \PYG{n}{para}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{+} \PYG{n}{in}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{]);}
+    \PYG{k}{return} \PYG{n}{cost}\PYG{p}{;}
+  \PYG{p}{\PYGZcb{}}
+
+  \PYG{k+kt}{void} \PYG{n}{Gradient}\PYG{p}{(}\PYG{k}{const} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{para}\PYG{p}{,} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{s}\PYG{p}{,} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{g}\PYG{p}{,} \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{bs}\PYG{p}{)} \PYG{p}{\PYGZob{}}
+    \PYG{n}{g}\PYG{p}{.}\PYG{n}{zeros}\PYG{p}{(}\PYG{n}{para}\PYG{p}{.}\PYG{n}{n\PYGZus{}rows}\PYG{p}{,} \PYG{n}{para}\PYG{p}{.}\PYG{n}{n\PYGZus{}cols}\PYG{p}{);}
+    \PYG{k}{for} \PYG{p}{(}\PYG{k+kt}{size\PYGZus{}t} \PYG{n}{i} \PYG{o}{=} \PYG{n}{s}\PYG{p}{;} \PYG{n}{i} \PYG{o}{\PYGZlt{}} \PYG{n}{s} \PYG{o}{+} \PYG{n}{bs}\PYG{p}{;} \PYG{n}{i}\PYG{o}{++}\PYG{p}{)}
+      \PYG{n}{g}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{+=} \PYG{p}{(}\PYG{l+m+mf}{1.0} \PYG{o}{/} \PYG{n}{bs}\PYG{p}{)} \PYG{o}{*} \PYG{l+m+mi}{2} \PYG{o}{*} \PYG{n}{para}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{])} \PYG{o}{+} \PYG{n}{bi}\PYG{p}{(}\PYG{n}{ord}\PYG{p}{[}\PYG{n}{i}\PYG{p}{]);}
+  \PYG{p}{\PYGZcb{}}
+
+  \PYG{n}{arma}\PYG{o}{::}\PYG{n}{vec} \PYG{n}{in} \PYG{c+cm}{/* intercepts */}\PYG{p}{,} \PYG{n}{bi} \PYG{c+cm}{/* coeffs */}\PYG{p}{;} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{uvec} \PYG{n}{ord} \PYG{c+cm}{/* function order */}\PYG{p}{;}
+\PYG{p}{\PYGZcb{};}
+\end{Verbatim}
 }
-
-  size_t NumFunctions() { return 4; }
-  void Shuffle() { ord = arma::shuffle(arma::uvec("0 1 2 3")); }
-
-  double Evaluate(const arma::mat& para, size_t s, size_t bs) {
-    double cost = 0;
-    for (size_t i = s; i < s + bs; i++)
-      cost += para(ord[i]) * para(ord[i]) + bi(ord[i]) * para(ord[i]) + in(ord[i]);
-    return cost;
-  }
-
-  void Gradient(const arma::mat& para, size_t s, arma::mat& g, size_t bs) {
-    g.zeros(para.n_rows, para.n_cols);
-    for (size_t i = s; i < s + bs; i++)
-      g(ord[i]) += (1.0 / bs) * 2 * para(ord[i]) + bi(ord[i]);
-  }
-
-  arma::vec in /* intercepts */, bi /* coeffs */; arma::uvec ord /* function order */;
-};
-\end{minted}
-\vspace*{-0.3em}
+\vspace*{-0.5em}
 
@@
 
-\vspace{-0.3em}
+\vspace{-0.5em}
 %\caption{Simple objective function where each dimension has a parabola with a
@@
 %\begin{listing}[H]
-\begin{minted}[fontsize=\footnotesize]{c++}
-struct SimpleOptimizer {
-  SimpleOptimizer(size_t bs = 1, double lr = 0.02) : bs(bs), lr(lr) { }
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+\PYG{k}{struct} \PYG{n}{SimpleOptimizer} \PYG{p}{\PYGZob{}}
+  \PYG{n}{SimpleOptimizer}\PYG{p}{(}\PYG{k+kt}{size\PYGZus{}t} \PYG{n}{bs} \PYG{o}{=} \PYG{l+m+mi}{1}\PYG{p}{,} \PYG{k+kt}{double} \PYG{n}{lr} \PYG{o}{=} \PYG{l+m+mf}{0.02}\PYG{p}{)} \PYG{o}{:} \PYG{n}{bs}\PYG{p}{(}\PYG{n}{bs}\PYG{p}{),} \PYG{n}{lr}\PYG{p}{(}\PYG{n}{lr}\PYG{p}{)} \PYG{p}{\PYGZob{}} \PYG{p}{\PYGZcb{}}
 
-  template<typename FunctionType>
-  double Optimize(FunctionType& function, arma::mat& parameter) {
-    arma::mat gradient;
-    for (size_t i = 0; i < 5000; i += bs) {
-      if (i % function.NumFunctions() == 0) { function.Shuffle(); }
-      function.Gradient(parameter, i % function.NumFunctions(), gradient, bs);
-      parameter -= lr * gradient;
+  \PYG{k}{template}\PYG{o}{\PYGZlt{}}\PYG{k}{typename} \PYG{n}{FunctionType}\PYG{o}{\PYGZgt{}}
+  \PYG{k+kt}{double} \PYG{n}{Optimize}\PYG{p}{(}\PYG{n}{FunctionType}\PYG{o}{\PYGZam{}} \PYG{n}{function}\PYG{p}{,} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat}\PYG{o}{\PYGZam{}} \PYG{n}{parameter}\PYG{p}{)} \PYG{p}{\PYGZob{}}
+    \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat} \PYG{n}{gradient}\PYG{p}{;}
+    \PYG{k}{for} \PYG{p}{(}\PYG{k+kt}{size\PYGZus{}t} \PYG{n}{i} \PYG{o}{=} \PYG{l+m+mi}{0}\PYG{p}{;} \PYG{n}{i} \PYG{o}{\PYGZlt{}} \PYG{l+m+mi}{5000}\PYG{p}{;} \PYG{n}{i} \PYG{o}{+=} \PYG{n}{bs}\PYG{p}{)} \PYG{p}{\PYGZob{}}
+      \PYG{k}{if} \PYG{p}{(}\PYG{n}{i} \PYG{o}{\PYGZpc{}} \PYG{n}{function}\PYG{p}{.}\PYG{n}{NumFunctions}\PYG{p}{()} \PYG{o}{==} \PYG{l+m+mi}{0}\PYG{p}{)} \PYG{p}{\PYGZob{}} \PYG{n}{function}\PYG{p}{.}\PYG{n}{Shuffle}\PYG{p}{();} \PYG{p}{\PYGZcb{}}
+      \PYG{n}{function}\PYG{p}{.}\PYG{n}{Gradient}\PYG{p}{(}\PYG{n}{parameter}\PYG{p}{,} \PYG{n}{i} \PYG{o}{\PYGZpc{}} \PYG{n}{function}\PYG{p}{.}\PYG{n}{NumFunctions}\PYG{p}{(),} \PYG{n}{gradient}\PYG{p}{,} \PYG{n}{bs}\PYG{p}{);}
+      \PYG{n}{parameter} \PYG{o}{\PYGZhy{}=} \PYG{n}{lr} \PYG{o}{*} \PYG{n}{gradient}\PYG{p}{;}
+    \PYG{p}{\PYGZcb{}}
+    \PYG{k}{return} \PYG{n}{function}\PYG{p}{.}\PYG{n}{Evaluate}\PYG{p}{(}\PYG{n}{parameter}\PYG{p}{,} \PYG{l+m+mi}{0}\PYG{p}{,} \PYG{n}{function}\PYG{p}{.}\PYG{n}{NumFunctions}\PYG{p}{());}
+  \PYG{p}{\PYGZcb{}}
+  \PYG{k+kt}{size\PYGZus{}t} \PYG{n}{bs}\PYG{p}{;} \PYG{k+kt}{double} \PYG{n}{lr}\PYG{p}{;} \PYG{c+c1}{// Store batch size and learning rate internally.}
+\PYG{p}{\PYGZcb{};}
+\end{Verbatim}
 }
-    return function.Evaluate(parameter, 0, function.NumFunctions());
-  }
-  size_t bs; double lr; // Store batch size and learning rate internally.
-};
-\end{minted}
-\vspace{-0.3em}
+\vspace{-0.5em}
 %\caption{Simple stochastic gradient descent method with a static number of
@@
 \vspace*{-0.4em}
-\begin{minted}[fontsize=\footnotesize]{c++}
-  ObjectiveFunction function; arma::mat parameter("0 0 0 0;");
-  SimpleOptimizer optimizer;
-  std::cout << "objective: " << optimizer.Optimize(function, parameter);
-\end{minted}
+{\footnotesize
+\begin{Verbatim}[commandchars=\\\{\}]
+  \PYG{n}{ObjectiveFunction} \PYG{n}{function}\PYG{p}{;} \PYG{n}{arma}\PYG{o}{::}\PYG{n}{mat} \PYG{n}{parameter}\PYG{p}{(}\PYG{l+s}{\PYGZdq{}0 0 0 0;\PYGZdq{}}\PYG{p}{);}
+  \PYG{n}{SimpleOptimizer} \PYG{n}{optimizer}\PYG{p}{;}
+  \PYG{n}{std}\PYG{o}{::}\PYG{n}{cout} \PYG{o}{\PYGZlt{}\PYGZlt{}} \PYG{l+s}{\PYGZdq{}objective: \PYGZdq{}} \PYG{o}{\PYGZlt{}\PYGZlt{}} \PYG{n}{optimizer}\PYG{p}{.}\PYG{n}{Optimize}\PYG{p}{(}\PYG{n}{function}\PYG{p}{,} \PYG{n}{parameter}\PYG{p}{);}
+\end{Verbatim}
+}
 \vspace*{-0.4em}
@@
 there would be when providing this type of support through inheritance or in
-other languages such as Python or C.
+languages such as Python or C.
 
